### 2013-05-06 koenker-zeileis-09/JAE-RER.tex
using sweave

\documentclass[10pt,a4paper,twoside]{article}

%% additional packages
\usepackage[latin1]{inputenc}
\usepackage{a4wide,graphicx,color,thumbpdf}
\usepackage{hyperref}
\usepackage{amsmath}

%% BibTeX settings
\usepackage[authoryear,round]{natbib}
\bibliographystyle{jae}
\bibpunct{(}{)}{,}{a}{,}{,}
\newcommand{\doi}[1]{\href{http://dx.doi.org/#1}{\normalfont\texttt{doi:#1}}}

%% markup commands for code/software
\let\code=\texttt
\let\pkg=\textbf
\let\proglang=\textsf
\newcommand{\file}[1]{`\code{#1}'}
\newcommand{\email}[1]{\href{mailto:#1}{\normalfont\texttt{#1}}}

%% paragraph formatting
\renewcommand{\baselinestretch}{1}

%% \usepackage{Sweave} is essentially
\RequirePackage[T1]{fontenc}
\RequirePackage{ae,fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}

%% mimic JAE style
\renewcommand{\section}{\secdef \mysec \mysecnn}
\newcommand{\mysec}[2][default]{\vspace{1.7\baselineskip}%
  \pdfbookmark[1]{#1}{Section.#1}%
  \refstepcounter{section}%
  \centerline{\large \thesection. \uppercase{#1}} \vspace{.5\baselineskip}}
\newcommand{\mysecnn}[1]{\vspace{1.7\baselineskip}%
  \centerline{\large #1} \vspace{.5\baselineskip}}
\renewcommand{\abstractname}{\normalfont SUMMARY}
\renewcommand{\refname}{REFERENCES}
\renewcommand{\thetable}{\Roman{table}}

%% page header (and currently no footer)
\usepackage{fancyhdr}
\setlength{\headheight}{15pt}
\renewcommand{\headrulewidth}{0pt}
\pagestyle{fancy}
\thispagestyle{plain}
\fancyhf{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[CE]{{\normalsize \uppercase{R.~Koenker and A.~Zeileis}}}
\fancyhead[CO]{{\normalsize \uppercase{On Reproducible Econometric Research}}} 
\fancyfoot[LO,LE]{\small Copyright {\copyright} 2009 John Wiley \& Sons, Ltd.}
\fancypagestyle{plain}{
\fancyhf{}
\fancyfoot[LO,LE]{\small This is a preprint of an article published in %
  \textit{Journal of Applied Econometrics}, \textbf{24}(5), 833--847. \\%
  Copyright {\copyright} 2009 John Wiley \& Sons, Ltd. \doi{10.1002/jae.1083} }
}

%% title information
\title{\bf On Reproducible Econometric Research}
\author{\hfill Roger Koenker$^a$ \hfill Achim Zeileis$^b$\thanks{Correspondence
to: Achim Zeileis, Department of Statistics and Mathematics, WU Wirtschaftsuniversit\"at Wien,
Augasse 2--6, 1090 Wien, Austria. Tel: +43/1/31336-5053. Fax: +43/1/31336-774.
E-mail: \email{Achim.Zeileis@R-project.org}} \hfill \hfill \\
{\small \it $^a$ Department of Economics, University of Illinois at Urbana-Champaign, United States of America} \\
{\small \it $^b$ Department of Statistics and Mathematics, WU Wirtschaftsuniversit\"at Wien, Austria}}
\date{}

% hyperref setup
\definecolor{Red}{rgb}{0.5,0,0}
\definecolor{Blue}{rgb}{0,0,0.5}
\hypersetup{%
  pdftitle = {On Reproducible Econometric Research},
  pdfsubject = {},
  pdfkeywords = {reproducibility, replication, software, literate programming},
  pdfauthor = {Roger Koenker, Achim Zeileis},
  %% change colorlinks to false for pretty printing
  colorlinks = {true},
  linkcolor = {Blue},
  citecolor = {Blue},
  urlcolor = {Red},
  hyperindex = {true},
  linktocpage = {true},
}

\begin{document}

\maketitle

\begin{abstract}
Recent software developments are reviewed from the vantage point of reproducible
econometric research.  We argue that the emergence of new tools, particularly in
the open-source community, have greatly eased the burden of documenting and archiving
both empirical and simulation work in econometrics.  Some of these tools are highlighted
in the discussion of two small replication exercises.
\end{abstract}

\noindent {\bf Keywords:} reproducibility, replication, software,
  literate programming.

\section{Introduction} \label{sec:intro}

The renowned dispute between Gauss and Legendre over priority for the
invention of the method of least squares might have been resolved by
\cite{repro:Stigler:1981}.  A calculation of the earth's ellipticity
reported by Gauss in 1799 alluded to the use of {\it meine Methode};
had Stigler been able to show that Gauss's estimate  was consistent
with the least squares solution using the four observations available
to Gauss, his claim that he had been using the method since 1795 would
have been strongly vindicated.  Unfortunately, no such computation
could be reproduced leaving the dispute in that limbo all too familiar
in the computational sciences.

The question that we would like to address here is this:  200
years after Gauss, can we do better?  What can be done to
improve our ability to reproduce computational results in econometrics
and related fields?  Our main contention is that recent software developments, 
notably in the open-source community, make it much easier to achieve 
and distribute reproducible research.

What do we mean by reproducible
research?  \cite{repro:Buckheit+Donoho:1995} have defined what
\cite{repro:deLeeuw:2001} has called \emph{Claerbout's Principle}:
``An article about computational science in a scientific publication
is \emph{not} the scholarship itself, it is merely \emph{advertising}
of the scholarship. The actual scholarship is the complete software
development environment and the complete set of instructions which
generated the figures.''  We view this as a desirable objective for
econometric research.  See \cite{repro:Schwab+Karrenbach+Claerbout:2000} 
for further elaboration of this viewpoint.

The transition of econometrics from a handicraft industry
\citep{repro:Wilson:1973,repro:Goldberger:2004}
to the modern sweatshop of globally interconnected computers
has been a boon to productivity and innovation, but sometimes
seems to be a curse.  Who among us expected to be in the ``software
development'' business?  And yet many of us find ourselves  precisely
in this position, and those who are not, probably should be.  As
we will argue below, software development is no longer something  that
should be left to specialized commercial developers, but instead should
be an integral part of the artisanal econometric research process.  Effective
communication of research depends crucially on documentation and
distribution of related software and data.

Some journals, such as the \emph{Journal of Applied Econometrics}
(JAE, \url{http://JAE.Wiley.com/}), support authors in this task
by providing data archives \citep{repro:MacKinnon:2007}. However,
a more integrated approach encompassing data, software, empirical
analysis, and documentation is often desirable to facilitate
replication of results.

\section{Software Tools} \label{sec:software}

In this section, we review some recent software developments that facilitate a more 
reproducible approach to econometric research. The tools discussed encompass
the most common components of econometric practice: from data handling, over
data analysis in some programming environment, to preparing a document describing
the results of the analysis. Additionally, we provide information on literate
programming tools and techniques that enable an integrated approach to these
three steps, and on software for version control of all components.

\subsection{Version Control} \label{sec:version}

Econometric research on a given project is often carried out over an extended
period, by several authors, on several computers, so it is hardly surprising
that we often have difficulty reconstructing exactly what was done.  Files
proliferate with inconsistent naming conventions, get overwritten or deleted
or are ultimately archived in a jumble with a sigh of relief when papers are finally
accepted for publication.  Sometimes, as is the case with JAE papers, a part of the 
archive is submitted to a central repository, but more often the archive resides
peacefully on one or more authors' computers until the next disk crash or
system upgrade.

Such difficulties have long been recognized in software development
efforts, but it is only relatively recently that practical version control systems
have emerged that are well adapted to econometric research.  Many
systems have grown out of the open-source software community where rigorous
archive cataloging of development efforts involving many participants is
absolutely essential.  We briefly describe one such system, \pkg{Subversion} 
\citep[SVN, \url{http://Subversion.Tigris.org/}, see][]{repro:Pilato+Collins-Sussman+Fitzpatrick:2004}
that we have used for this project.
Cross-platform compatibility is an important consideration in many projects;
SVN's basic command-line interface will feel comfortable to most 
Linux/Unix users, but the graphical clients \pkg{TortoiseSVN}
(\url{http://TortoiseSVN.Tigris.org/}), embedded 
into the Windows \pkg{Explorer}, or  \pkg{svnX} for 
Mac (\url{http://www.Apple.com/downloads/macosx/development_tools/svnx.html})
will prove convenient for others.

The first stage of any SVN project involves the creation of a central repository
for the project.  Once  the repository is created, any of the authorized collaborators
can ``checkout'' the current version of the project files, update the local copies
of the files, and make changes, additions or deletions to the files.  At any point,
the modified  file structure  can be ``committed'' thereby creating a new version
of the project.  When changes are committed they are recorded in the repository
as incremental modifications so that all prior versions are still fully accessible
(even if a file was deleted from the current revision).
In rare cases that more than one author has modified the same passage of the same
file, the system prompts authors to reconcile the changes before accepting the
commitment.  A complete historical record of the project is available at any point in 
the development process for inspection, search and possible restoration.
Since modifications are stored as increments, file ``diffs'' in Unix jargon,
storage requirements for the repository are generally far less severe than for
prior improvised archiving strategies.  Version numbering is fully automatic,
so one consistent set of naming conventions for project files can be adhered
to, thereby avoiding the typical m\'elange of ad hoc files resulting from impromptu
version control strategies.

If an SVN server is available, setting up an SVN project/repository is very easy.
Only if no SVN server is available, the setup requires some technical work.
It is not dissimilar to setting up a Web server but can be a lot easier
\citep[see][Chapter~6]{repro:Pilato+Collins-Sussman+Fitzpatrick:2004}. Also,
the individual researcher can often avoid setting up a server alone, especially if the university/department
already provides such a server, typically along with other Web services.
Some Web pages also host free SVN repositories, esepecially for software projects, e.g.,
\url{http://SourceForge.net/} or \url{http://Code.Google.com/},
and \url{http://R-Forge.R-project.org/} among others.

\subsection{Data Technologies and Data Archiving} \label{sec:data}

%% 1. flat data
In the beginning data was flat, representable as matrices, and easily storable in
text files.  Easily stored need not imply easily retrieved, as anyone with 
with a drawer full of floppy disks or a collection of magnetic tapes 
is undoubtedly aware. As data has become more plentiful it has also become
more structurally complex.
%% 2. DBMS
Relational database management systems (DBMSs) are still
quite exotic in econometrics, but seem likely to become more commonplace.
Stable, well-documented databases with proper version control are critical to
the reproducibility goal. 
%% 3. Web-based formats
In many cases, data archives for individual projects can be most efficiently
represented by software describing selection rules from publicly available
data sources, like the Panel Study on Income Dynamics (PSID) and the US
Census Bureau. Such data sources often provide their contents via the World
Wide Web (WWW), either using one of the data technologies above or new standards
such as \proglang{XML} or \proglang{PHP} 
that emerged with the WWW.
However, as we stress in Section~\ref{sec:growth} regarding the Penn World Tables,
reproducibility of such data retrieval strategies relies heavily on future access to
current versions of the database.

%% 1. flat data
If the data to be stored is not too complex and/or large, a flat plain text file
is often ideal for reproducibility (and hence used for most publications in the
JAE data archive): it is portable across platforms, many software
packages can read and write text files and they can be efficiently archived, e.g.,
in an SVN repository.
For larger data sets, it may be useful or even necessary to store the data
in a DBMS. Many candidates are available here, but most use
some form of the \proglang{SQL} (structured query language).
Two popular open-source candidates, available across many platforms, are
\pkg{PostgreSQL} (\url{http://www.PostgreSQL.org/}) and \pkg{MySQL} (\url{http://MySQL.com/}).

%% 3. Web-based formats
In addition to the data technologies above, other new standards have emerged,
especially for sharing data across the WWW. These include 
\proglang{XML} \citep[extensible markup language,][]
{repro:Bray+Paoli+SperbergMcQueen:2008}: a text-based format,
well-suited for storing both data and meta-information, which is therefore used
as the basis for many other data and text formats. A key feature of \proglang{XML}
is that it is an open standard, maintained by an international consortium,
in contrast to many proprietary formats which are often altered periodically.
Data formats based on open standards have a higher likelihood of remaining accessible in the future. 
Further discussion of data technologies for statistical/scientific applications may
be found in \cite{repro:Murrell:2009}.

\subsection{Programming Environments} \label{sec:programming}

Econometrics has always been a Tower of Babel with many languages, or programming
environments,  
competing for attention in the never-ending struggle to communicate effectively
with machines. Over the course of our disparate careers we have had
opportunities to explore many of these, including, in approximate chronological order:  
\proglang{TSP}, \proglang{MIDAS}, \proglang{SAS}, \proglang{Statlib},
\proglang{GLIM}, \proglang{S}, \proglang{LIMDEP}, \proglang{SHAZAM}, \proglang{GAUSS},
\proglang{S-PLUS}, \proglang{Lisp-Stat}, \proglang{SPSS}, \proglang{Minitab},
\proglang{Stata}, \proglang{Ox}, \proglang{MATLAB}, \proglang{Mathematica}, and 
\proglang{R}. Clearly, these vary considerably in their level of ``programmability'' and
degree of specialization. Lower-level languages such as \proglang{C} or \proglang{Fortran}
also play an important role. A historical survey of the early development of econometric software
may be found in \cite{repro:Renfro:2004}, see also \cite{repro:Ooms+Doornik:2006}. 
There is also an extensive literature on the individual merits and comparative 
performance of various forms of statistical software;  see \cite{repro:Baiocchi:2007} 
for further references, and a more proscriptive review of reproducibility issues.
Here, we will try to restrict our attention quite narrowly to properties of
programming environments that might facilitate reproducible research.

\paragraph{Language Structure}
%% 1. readability
Software reviews of programming environments often stress speed
of execution, rather than accuracy, breadth of coverage, or other less tangible
attributes. This is unfortunate since the exertions of the machine are rarely a
matter concern, or sympathy. Of more direct relevance in our view is ease of
writing software (for avoiding errors) and reading software
(for replication of results).  For those who believe that computer
programming should be, like prayer, a private language between the individual
supplicant and the \emph{deus ex machina}, we recommend 
\cite{repro:Kernighan+Plauger:1974} who emphasize the need to write programs 
that other \emph{people} are capable of reading.  The programming language
should support the user in turning theory into software that reflects how
we think about the underlying method conceptually. Several language features are 
helpful in obtaining this goal: functional languages, object orientation, and
modular programs with re-usable components.
%% functioncal languages
Environments for statistical
computing are gradually moving away from the ``canned soup model'' toward 
functional languages  that permit fresh ingredients to be assembled in a more flexible
manner. Formulae specification for linear and nonlinear models, and unified  
interfaces for general linear models constitute two developments that have
brought software and theoretical specification of models closer together.  
%% object orientation
Combining such an approach with object orientation allows one to encapsulate
complex structures (such as fitted regression models) and define typical
tasks for them such as inference or visualization. Programs or analyses
written in such a way are more intelligible and hence easier to reproduce.
%% re-usability
Furthermore, it is highly desirable to have a single environment within which
to conduct empirical analyses and simulation studies. Re-using the same functionality 
across different tasks assures better consistency and avoids duplication of programming effort.
But environments designed for the convenience of empirical analysis may not provide
good simulation environments, and vice-versa.  To assure reproducibility
and re-usability by other authors, the structural features of a language should
facilitate (and not suppress) the ability to build on innovations of prior authors.  
%% further comments
Exploiting common structure in problems
often leads to general software solutions that facilitate critical comparison
of various methods as illustrated for structural change testing
in \cite{repro:Zeileis:2006a}. 
The modern trend toward functional
languages and object-orientation is visible in several of the currently 
successful econometrics environments.
\proglang{MATLAB} (\url{http://www.MathWorks.com/}),
\proglang{Ox} (\url{http://www.doornik.com/ox/}) and
\proglang{R} (\url{http://www.R-project.org/})
are notable in this respect.

\paragraph{Open Sources}
Languages are acquired by mimicry so it is extremely valuable to have access to a
large body of text written by those proficient in the language.  This is one of
the prime advantages of open-source software projects. 
When we first embark on a new econometric project, we have at our fingertips an 
extensive body of existing code, some of which will be directly useful as building 
blocks---provided that the language is structured to facilitate such modularity---and
some code will offer only stylistic hints.  In either case, adapting prior code
when available from reliable sources is almost always preferable to reinventing
the wheel---and, as argued above, increases the readability and intelligibility 
of the resulting programs. Most languages now offer some sort of forum for sharing code,
and thus provide access to an existing body of tested code. An
exemplary approach is the Comprehensive \proglang{R} Archive Network
(CRAN, \url{http://CRAN.R-project.org/}) that hosts about 1,700 software
packages (containing code, data, and documentation) which are checked on 
a daily basis on several platforms.
In addition to such open platforms for user-contributed code,
there are also several journals publishing peer-reviewed software, e.g.,
\emph{The Stata Journal} (\url{http://www.Stata-Journal.com/}) and the
\emph{Journal of Statistical Software} (\url{http://www.JStatSoft.org/}).
Furthermore, even commercial software producers usually provide some functionality written in
the language itself and therefore open to inspection and emulation for the individual user.  
\proglang{MATLAB} and \proglang{Stata} (\url{http://www.Stata.com/})
are notable in this respect. However, even in these favorable circumstances 
one may eventually wish to dig below the surface only to discover that crucial
elements of the story are accessible only in a binary form readable only
by machines.  At this point the computations become a black box visible only
to the select few, and scientific trust becomes a leap of faith.
Source code is itself the ultimate form of documentation for computational
science. When it is well written it can serve as an inspiration for subsequent
work. When it is flawed, but accessible, then it can be much more easily 
subjected to necessary critical scrutiny.
With computational methods, gradual refinement is seriously impeded
unless source code is open.

\paragraph{Language Interfaces}
To combine the convenience of high-level programming environments with 
the efficiency of compiled languages, it is desirable to either (byte-)compile
program code directly or to be able to link code written in lower-level
languages like \proglang{Fortran} or \proglang{C} and their dialects.
Most modern languages employed in econometrics---including \proglang{MATLAB},
\proglang{Ox}, \proglang{R}, and \proglang{Stata}---offer some means
for ordinary users to accomplish this sort of sorcery. Doing so in ways that
are platform independent is a considerable challenge. From a reproducibility
perspective, it should be assured that even code interfacing lower level languages
behaves consistently across platforms with differing hardware and operating systems.
This problem is particularly acute in simulation settings where it is often
desirable to be able to reproduce sequences pseudo-random numbers across machines
and operating systems.

\paragraph{Environmental Stability}
Since hardware and software environments for econometric research are constantly
evolving, a major challenge for reproducibility involves proper documentation of 
the environment used to conduct the computational component of the project.  Ideally,
in our view, authors and journals should be expected to provide a section similar to
Section~\ref{sec:computational} (``Computational Details'') of the present paper
describing in some detail the software and hardware environment employed. 
Even when these environments are completely specified, it is likely to be difficult
several years later to restore
a prior version of the environment.  In this respect, again, the open-source
community is exemplary, since prior versions are typically archived and easily
downloadable. For commercial software prior versions are sometimes difficult to obtain due to
licensing and distribution constraints. 
\cite{repro:Zeileis+Kleiber:2005} describe 
an exercise in ``forensic econometrics'' exploring the evolution of the evaluation
of the Gaussian distribution function in successive versions of
\proglang{GAUSS}. 
Such investigations are obviously handicapped by the unavailability of older versions
and lack of adequate documentation of algorithms and their evolution.

\paragraph{User Interfaces}
In the paragraphs above, we have focused on reproducibility using source code
as typically used in programs with a command line interface (CLI). In practice,
however, many analyses are carried out using a point-and-click approach
based on a graphical user interface (GUI). Although easier to learn and often
easier to apply, this procedure poses a challenge to reproducibility because
it is much harder to reconstruct the ``click history'' for a certain analysis.
A useful compromise are log files offered by several GUI-based systems that
contain the source code associated with the point-and-click analysis. Thus,
an author who wants to assure reproducibility, probably cannot avoid the contact
with source code alltogether, but the transition to reading/writing source
code is often made easier in this way.

\subsection{Document Preparation Systems} \label{sec:document}

{\TeX} and {\LaTeX} have become the \emph{lingua franca} for the composition 
of mathematical text in general and econometric text in particular.
{\TeX} (\url{http://www.TUG.org/}) developed by \cite{repro:Knuth:1984}
beginning in the late 1970s constitutes an exceptional case study in 
software development and the effectiveness of the gradual refinement process 
of open-source projects.  
{\LaTeX} (\url{http://www.LaTeX-project.org/}), originally written by  \cite{repro:Lamport:1994},
still constitutes an ongoing development effort to build a higher level markup
language on the foundation provided by {\TeX}.  Nevertheless, {\LaTeX} is also
a remarkably stable environment and serves as a convenient and portable format
across many platforms for a wide variety of documents.

Although these systems are clearly state of the art in econometric document
preparation, casual users of document preparation systems often prefer to avoid
the somewhat steeper learning curve of {\LaTeX} in favor of \emph{WYSIWYG}
(what you see is what you get) text processors such as Microsoft's \pkg{Word}
(\url{http://www.Office2007.com/}) or \pkg{OpenOffice.org} (\url{http://www.OpenOffice.org/}).
To avoid a general
discussion about the relative merits of {\LaTeX} over WYSIWYG processors, we focus
on the perspective of reproducibility: Especially \pkg{Word}'s proprietary binary document
format is problematic in this respect as its documents are less stable across platforms or
versions of \pkg{Word}. With the advent of the open-source suite \pkg{OpenOffice.org}
the situation improved considerably as it not only introduced an \proglang{XML}-based
open document file format (ODF) for WYSIWYG documents but also supports export to 
a number of older proprietary file formats.

\subsection{Literate Programming} \label{sec:literate}

The idea of merging text, documentation and various forms of computer code
arose from the structured programming discussions of the 1970s and was
championed by \cite{repro:Knuth:1992} following his initial development of
{\TeX}.  Literate programming, as this movement has come to be called, 
encourages a more readable programming style by making the code itself an
integral part of the documentation. The basic operations on documents 
containing both code chunks and documentation chunks are known as \emph{tangling} and
\emph{weaving}: the former strips off the documentation chunks and extracts only
the code chunks while the latter weaves the code chunks into the documentation,
typically by adding appropriate markup for display of the code. Following
\cite{repro:Knuth:1992}, the initial literate programming systems were
\pkg{WEB} and \pkg{CWEB} for combining \proglang{Pascal} and \proglang{C}
code, respectively, with {\TeX} documentation. In order to provide a leaner
and more flexible literate programming system, \cite{repro:Ramsey:1994}
developed \pkg{noweb}, a set of open-source tools for combining code in
arbitrary languages and {\LaTeX} documentation.

Literate programming is an important first step in supporting reproducibility in
econometric practice but one would like to carry the idea a step further and include not only
the code but also its results dynamically in a document \citep{repro:Leisch+Rossini:2003}. 
This idea of \emph{literate econometric practice} seems especially well-suited to statistical
computing applications where models, data, algorithms and interpretation all coalesce
to produce scientific documents. Directly linking text with computational procedures
reduces the likelihood of inconsistencies and facilitates reproducing the same type of
analysis with an extentended/modified data set or different parameter
settings.  Proximity is, of course, no guarantee that text and code will agree; we have 
all read and probably written commented code for which the comments contradicted the 
unintended consequences of the code. 
\proglang{Mathematica}'s concept of ``notebooks'' is an approach to documents of this type
in which the displayed document can be easily altered by changing the associated \proglang{Mathematica}
code. Another implementation, more closely modelled after the literate programming ideas
discussed above, is the \code{Sweave} system of \cite{repro:Leisch:2002} for \proglang{R}.
Re-using the markup commands of \pkg{noweb}, it allows authors to create ``revivable'' documents
containing a tightly coupled bundle of code and documentation. To illustrate this,
and to try to practice what we preach, the archived version of this paper includes
a source file \file{JAE-RER.Rnw}
that includes all of the text as well as all of the code used to generate
the statistical analyses presented in the next section. The input file is structured
into code chunks written in \proglang{R} and text chunks written in {\LaTeX} and it
can be processed in \proglang{R} with the command \verb|Sweave("JAE-RER.Rnw")|.
This extracts the code chunks and executes them in \proglang{R}, produces the associated
output (either in numerical or graphical form) and weaves them into a 
{\LaTeX} file. This resulting {\LaTeX} file can
then be processed to PDF by pdf{\LaTeX}. The choice of pdf{\LaTeX} is specific to our document.
Many other flavours of \code{Sweave} are available, including {\LaTeX},
HTML \citep[\pkg{R2HTML},][]{repro:Lecoutre:2003}, and ODF \citep[\pkg{odfWeave},][]{repro:Kuhn:2006}.
There are also extensions to mixing \proglang{SAS} code with {\LaTeX} documentation
\citep{repro:Lenth+Hojsgaard:2007}.

\section{Replication Case Studies} \label{sec:replication}

In this section, we describe briefly two replication exercises chosen
to illustrate several aspects of the reproducibility problems discussed above.
The software tools used are \pkg{Subversion} for version control, flat text files
for data storage, \proglang{R} \citep{repro:R:2007} for programming and empirical analysis, and {\LaTeX}
for document preparation. As mentioned above, a literate data analysis approach is
adopted based on the \code{Sweave} tools. The SVN archive, containing the
full sources of the document, can be checked out anonymously from
\url{svn://svn-statmath.wu-wien.ac.at/repro/}. For convenience of non-SVN users,
there is also an online archive available at \url{http://www.econ.uiuc.edu/~roger/research/repro/}.
An extended version \citep{repro:Koenker+Zeileis:2007} with many more details is also available
from the same locations.

\subsection{An Empirical Example: Cross-Country Growth Regression} \label{sec:growth}

\citet[henceforth DJ]{repro:Durlauf+Johnson:1995} investigate cross-country
growth behavior based on an extended Solow model suggested in
\citet[henceforth MRW]{repro:Mankiw+Romer+Weil:1992}. The variables in the growth regression
model are real GDP per member of working-age population, $Y/L$
(separately for 1960 and 1985); fraction of real GDP devoted to investment, $I/Y$
(annual average 1960--1985); growth rate of working-age population, $n$ (annual average 1960--1985);
fraction of working-age population enrolled in secondary school, $\mathit{SCHOOL}$
(annual average 1960--1985); and the adult literacy rate, $\mathit{LR}$ in 1960.
Data for these variables (except $\mathit{LR}$) for 121 countries is printed
in MRW and provided in electronic form in the JAE data archive along
with DJ (who added $\mathit{LR}$).

The unconstrained extended Solow model suggested by MRW in their Table~V and given by DJ in their
Equation~7 regresses GDP growth $\log(Y/L)_{1985} - \log(Y/L)_{1960}$ on initial GDP
$\log(Y/L)_{1960}$ as well as $\log(I/Y)$, $\log(n + g + \delta)$ (assuming $g + \delta = 0.05$),
and $\log(\mathit{SCHOOL})$. DJ first fit the model by least squares for all
98 non-oil-producing countries (reproducing the 
results of MRW) and to various subsets of countries---with subset selection based
on initial output $(Y/L)_{1960}$ and/or literacy $\mathit{LR}_{1960}$---finding multiple
regimes rather than a single stable set of coefficients.

Here, we aim to reproduce the unconstrained regression results given by DJ in their Tables~II and V
given the sample splits described in the paper. Although this seems to be a rather modest task, it turned
out to be surprisingly difficult, illustrating some typical pitfalls. We first read the data,
provided in file \file{data.dj} in the JAE data archive, into \proglang{R}, coding missing values
as described in the accompanying documentation and subsequently selecting the non-oil-producing
countries.
%
\begin{Schunk}
\begin{Sinput}
R> dj <- read.table("data.dj", header = TRUE, na.strings = c("-999.0", "-999.00"))
R> dj <- subset(dj, NONOIL == 1)
\end{Sinput}
\end{Schunk}
%
The relevant columns in the resulting data set \code{dj} are \code{GDP85},
\code{GDP60}, \code{IONY}, \code{POPGRO}, \code{SCHOOL}, and \code{LIT60}.
The last four are described as ratios/fractions, but it was not clear from the
documentation that they were given in percent. Of course, this is quickly revealed by
a look at the actual data; and MRW probably used this scaling because it is easier to
read when printed. However, it remains unclear which scaling of the variables
was used for model fitting. After first attempting to use the variables as printed, it
turned out that MRW had scaled them to the unit interval. Thus,
the model employed by MRW and DJ (Table~II, first column)
can be written in \proglang{R} as the formula
%
\begin{Schunk}
\begin{Sinput}
R> mrw_model <- I(log(GDP85) - log(GDP60)) ~ log(GDP60) + log(IONY/100) +
+    log(POPGRO/100 + 0.05) + log(SCHOOL/100)
\end{Sinput}
\end{Schunk}
%
This model can now be fit with \proglang{R}'s linear model
function \code{lm()} to the \code{dj} data set, producing the following regression
summary:
%
\begin{Schunk}
\begin{Sinput}
R> dj_mrw <- lm(mrw_model, data = dj)
R> summary(dj_mrw)
\end{Sinput}
\begin{Soutput}
Call:
lm(formula = mrw_model, data = dj)

Residuals:
    Min      1Q  Median      3Q     Max 
-0.9104 -0.1760  0.0179  0.1844  0.9385

Coefficients:
                       Estimate Std. Error t value Pr(>|t|)
(Intercept)              3.0215     0.8275    3.65  0.00043
log(GDP60)              -0.2884     0.0616   -4.68  9.6e-06
log(IONY/100)            0.5237     0.0869    6.03  3.3e-08
log(POPGRO/100 + 0.05)  -0.5057     0.2886   -1.75  0.08306
log(SCHOOL/100)          0.2311     0.0595    3.89  0.00019

Residual standard error: 0.327 on 93 degrees of freedom
Multiple R-squared: 0.485,	Adjusted R-squared: 0.463 
F-statistic: 21.9 on 4 and 93 DF,  p-value: 8.99e-13 
\end{Soutput}
\end{Schunk}
%
reproducing coefficient estimates, standard errors, $\bar R^2$, and the
residual standard error $\sigma_\varepsilon$ provided in the first column of Table~II 
in DJ (identical to the first column in Table~V of MRW) with only small deviations 
for some of the entries.

In the next step we want to fit the two other columns of DJ's Table~II, pertaining to
the same model fitted to two different subsets: DJ employ a low-output/low-literacy
sample of 42 countries with $(Y/L)_{1960} < 1950$ and $\mathit{LR}_{1960} < 54\%$ and
the corresponding high-output/high-literacy sample, also consisting of 42 countries. 
However, when selecting these sub-samples and computing their size we obtain
sample sizes of 43 and 39, respectively,
showing that either the subset definitions or the sample sizes are misstated in 
DJ's Table~II.  Some brute-force search (via all-subsets regression) coupled 
with some educated guessing, yielded revised 
cut-offs for model fitting of 1800 and 50\%,
respectively, yielding consistent sample sizes, 
and also consistent model fits as we will see below.

Fitting the same model to these two subsets yields DJ's slope
coefficients, but the wrong intercepts. After further combinatorial search
using different logarithm bases and variable scalings it turned out that DJ 
appear to have employed the model: 
%
\begin{Schunk}
\begin{Sinput}
R> dj_model <- I(log(GDP85) - log(GDP60)) ~ log(GDP60) + log(IONY) +
+    log(POPGRO/100 + 0.05) + log(SCHOOL)
\end{Sinput}
\end{Schunk}
%
i.e., only scaling \code{POPGRO} to the original unit interval but keeping \code{IONY}
and \code{SCHOOL} in percent. Thus, the intercepts in their Table~II are not comparable
between the first and the other two columns, while the coefficients of interest are
unaffected by the scaling. Using the model formula \code{dj\_model}, we can then go on
and fit the model to both sub-samples:
%
\begin{Schunk}
\begin{Sinput}
R> dj_sub1 <- lm(dj_model, data = dj, subset = GDP60 <  1800 & LIT60 <  50)
R> dj_sub2 <- lm(dj_model, data = dj, subset = GDP60 >= 1800 & LIT60 >= 50)
\end{Sinput}
\end{Schunk}
%
%
%
and perform the Wald test for all coefficients, via \code{coeftest(dj\_sub1, vcov = sandwich)},
which reproduces the results from the second column in their Table~II (with only
minor deviations). Note that heteroskedasticity-consistent sandwich
standard errors are used
\citep[provided by package \pkg{sandwich} in \proglang{R}, see][]{repro:Zeileis:2004,repro:Zeileis:2006}
whereas conventional standard errors are used for the first MRW column.
Results for \code{dj\_sub2} can be obtained analogously.
All \proglang{R} output is provided in condensed form
in our Table~\ref{tab:dj} (generated using \code{Sweave()}
with the models fitted above).

\begin{table}[t]
\caption{\label{tab:dj}
  Replication of cross-country growth regressions, corresponding to
  models \code{dj\_mrw}, \code{dj\_sub1}, \code{dj\_sub2} with dependent
  variable $\log(Y/L)_{1985} - \log(Y/L)_{1960}$. Conventional standard errors
  are used in the first column and sandwich standard errors in the other two.}
\begin{center}
\begin{tabular}{lrrr} \hline
& & $(Y/L)_{1960} < 1800$ & $(Y/L)_{1960} \ge 1800$ \\
& MRW & $\mathit{LR}_{1960} < 50\%$ & $\mathit{LR}_{1960} \ge 50\%$ \\ \hline
Constant & $3.022\phantom{)}$ & $1.400\phantom{)}$ & $0.450\phantom{)}$ \\
 & $(0.827)$ & $(1.846)$ & $(0.723)$ \\
$\log(Y/L)_{1960}$ & $-0.288\phantom{)}$ & $-0.444\phantom{)}$ & $-0.435\phantom{)}$ \\
 & $(0.062)$ & $(0.157)$ & $(0.085)$ \\
$\log(I/L)$ & $0.524\phantom{)}$ & $0.310\phantom{)}$ & $0.689\phantom{)}$ \\
 & $(0.087)$ & $(0.114)$ & $(0.170)$ \\
$\log(n + g + \delta)$ & $-0.506\phantom{)}$ & $-0.379\phantom{)}$ & $-0.545\phantom{)}$ \\
 & $(0.289)$ & $(0.468)$ & $(0.283)$ \\
$\log(\mathit{SCHOOL})$ & $0.231\phantom{)}$ & $0.209\phantom{)}$ & $0.114\phantom{)}$ \\
 & $(0.059)$ & $(0.094)$ & $(0.164)$ \\ \hline
$\bar R^2$ & $0.46\phantom{0)}$ & $0.27\phantom{0)}$ & $0.48\phantom{0)}$ \\
$\sigma_\varepsilon$ & $0.33\phantom{0)}$ & $0.34\phantom{0)}$ & $0.30\phantom{0)}$ \\ \hline\end{tabular}
\end{center}
\end{table}

Thus, we have reproduced DJ's Table~II but only after some effort: cut-offs
for subset selection are different than those displayed in the table, variable scaling is different
leading to different intercepts, and the standard errors used vary across columns (although
DJ state briefly in a footnote that they would use sandwich standard errors). None of this
changes their results qualitatively: all conclusions drawn from the analysis of DJ are supported. 
Nevertheless, it would be desirable to avoid such uncertainty both from the author's
and reader's point of view. Organizing both code and documentation in a single revivable document
as we suggest in the previous section will assist in this but can, of
course, not prevent human error. It will, however, be much easier for the authors and readers---provided
the code is made publicly available---to find the sources of such confusions and untie the knots.

\subsection{Another Empirical Example: Wage-Equation Meta-Model} \label{sec:wage}

Our second case study in replication reports our experience trying to reproduce
the empirical results in one of our own papers.  \cite{repro:Koenker:1988} describes
a meta-analysis of published wage equations intended to explore how parametric
dimension of econometric models depends upon sample size.  
The data for the study consists of 733 wage equations from  
156 published papers
in mainstream journals and essay colllections.  In addition to citation information
and a topic indicator, the sample size and parametric dimension of the model was
recorded for each equation.

The models used for the analysis are standard count data models.  Observations 
on each wage equation are weighted by the reciprocal of the number of equations, $m$,
appearing in the paper, so the effective unit of analysis is really the paper not 
the equation.  The first, and simplest, version of the model, appears in the paper as
\begin{equation}
  \log \lambda \quad = \quad  \underset{(0.149)}{1.336} ~+~ \underset {(0.017)}{0.235} \log z.
\end{equation}
where $\lambda$ is Poisson rate parameter, the expected number of parameters $y$ in
the equation, and $z$ is the sample size.  The parameter of interest is the
elasticity of parsimony, or \emph{parsity}, the log derivative of $\lambda$ with respect
to $z$.   In this case, parsity is constant and equal to about $1/4$ implying
that parametric dimension increases roughly like the fourth root of the
sample size.  The paper dutifully reports that the computations were conducted
using release~3 of the \proglang{GLIM} system \citep{repro:Baker+Nelder:1978}.
Our first attempt to replicate these results in \proglang{R} invoked the incantation:
%
\begin{Schunk}
\begin{Sinput}
R> rk <- read.csv("rk.csv")
R> names(rk)[2:4] <- c("m", "y", "z")
R> rk1 <- glm(y ~ log(z), weights = 1/m, family = quasipoisson, data = rk)
R> coeftest(rk1)
\end{Sinput}
\begin{Soutput}
z test of coefficients:

            Estimate Std. Error z value Pr(>|z|)
(Intercept)   1.3362     0.1739    7.68  1.5e-14
log(z)        0.2353     0.0199   11.82  < 2e-16
\end{Soutput}
\end{Schunk}
%
%
The reader can imagine our relief seeing that the estimated coefficients
agreed perfectly with the published results, turning to dismay as we observed
that the standard errors differed by a factor of 
0.859.  
Covariance matrices for the parameters in quasi-Poisson models require 
an estimated dispersion parameter:
typically some generalization of a residual sum of squares scaled by the
residual degrees of freedom. In the text of the paper, this was claimed
to be the scaled Pearson statistic
$\hat \sigma_P^2 = (n-p)^{-1} \sum w \cdot (y - \hat y)^2 / \hat y$,
following the advice offered by \citet[pp.~172--173]{repro:McCullagh+Nelder:1983}.
Unfortunately, however, the authors of the \proglang{GLIM} system, in their wisdom,  
chose to use the corresponding \emph{deviance} statistic instead of the 
Pearson statistic, that is
they scaled the variance-covariance matrix of the estimated coefficients by
$\hat \sigma_D^2 = (n-p)^{-1} 2 \sum w \cdot (y \log(y/\hat y) - (y - \hat y))$.
\proglang{R} adopts the Pearson-based estimate as its default and hence the
dispersion of 6.42
reported in the summary above does not conform with the estimate of
4.73 reported in the 
original paper.  Recalculating the \proglang{R} results using the
deviance-based estimate
%
\begin{Schunk}
\begin{Sinput}
R> dispersion <- function(object, type = "deviance")
+    sum(residuals(object, type = type)^2)/df.residual(object)
R> summary(rk1, dispersion = dispersion(rk1))
\end{Sinput}
\end{Schunk}
%
reproduces the published results; see the first column of Table~\ref{tab:rk} for condensed output.
The lesson of this ``conflict of defaults'' is that it is dangerous  to make assumptions about
what software is doing; more careful reading of the \proglang{GLIM} manual would have
revealed that deviance residuals were being used and this error could have been
avoided.

In an effort to explore the sensitivity of the parsity estimate to the specification of the
functional form of the model, several other models were reported in 
\cite{repro:Koenker:1988}.  Table~\ref{tab:rk} summarizes our attempt to replicate
these results.  For each of the four Poisson models we report, in addition to
coefficients and standard errors, both the Pearson and deviance dispersion estimates
and an estimate of the parsity parameter  $\pi$ evaluated at $z = 1000$ which is
roughly the geometric mean of the observations on sample size.  The quadratic-in-logs
model was very sensitive to a few outlying $z$ observations and two versions of the
estimates were reported, one for the full sample as Equation~2 and one for a reduced
sample excluding points with $z > 250,000$, as Equation~3.  Again, replication 
revealed an error:  the text of the paper  reports that this cut-off was $500,000$,
but a yellowing page in the project file folder and the printed \proglang{GLIM}
output in the same folder clearly reveal that $250,000$ was used.  
This is a clear case in which a more
literate style of programming might have prevented the error by enforcing consistency
between the text and the econometric estimation.  Our use of \code{Sweave}  is one
way to accomplish this.
%
\begin{Schunk}
\begin{Sinput}
R> rk2 <- glm(y ~ log(z) + I(log(z)^2), weights = 1/m, family = quasipoisson,
+    data = rk)
R> rk3 <- glm(y ~ log(z) + I(log(z)^2), weights = 1/m, family = quasipoisson,
+    data = rk, subset = z <= 250000)
\end{Sinput}
\end{Schunk}
%
The results reported in Table~\ref{tab:rk}  for Equations~1 through 4 agree with the
published results to the precision reported, except for the coefficient on $\log z$
in Equation~2 which appears to be an old-fashioned typo (because the published parsity
estimate is consistent with our replicated coefficient rather than the published coefficient).
Mistakes like this could again be more easily avoided by better integration of text and data analysis.

Two final models employed $\log \log z$ as the only covariate; 
first in the quasi-Poisson specification (Equation~4) and then using the negative binomial
likelihood. The former can again be reproduced via
%
\begin{Schunk}
\begin{Sinput}
R> rk4 <- glm(y ~ log(log(z)), weights = 1/m, family = quasipoisson, data = rk)
\end{Sinput}
\end{Schunk}
%
using the same specification of dispersion as above. 
The negative binomial results in \cite{repro:Koenker:1988} were computed, not
in \proglang{GLIM}, but with general code written by Richard Spady for maximum
likelihood estimation and linked to \proglang{S}. The same model can now
be estimated in \proglang{R} using the \code{glm.nb()} function from the 
\pkg{MASS} package \citep{repro:Venables+Ripley:2002}.
%
\begin{Schunk}
\begin{Sinput}
R> rk_nb <- glm.nb(y ~ log(log(z)), data = rk, weights = 1/m)
\end{Sinput}
\end{Schunk}
%
%
%
\begin{table}[t]
\caption{\label{tab:rk}
  Replication of wage-equation meta-models, corresponding
  to \code{rk1}--\code{rk4} and \code{rk\_nb} with dependent variable $y$.  
  Deviance-based dispersion estimates are used for the quasi-Poisson models
  and sandwich standard errors for the negative binomial model.}
\begin{center}
\begin{tabular}{lrrrrr} \hline
& \multicolumn{4}{c}{Quasi-Poisson} & Neg-Bin \\
& \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{3}
& \multicolumn{1}{c}{4} & \\ \hline
Constant & $1.336\phantom{)}$ & $-0.439\phantom{)}$ & $1.737\phantom{)}$ & $-0.777\phantom{)}$ & $-0.677\phantom{)}$ \\
 & $(0.149)$ & $(0.512)$ & $(0.517)$ & $(0.315)$ & $(0.383)$ \\
$\log z$ & $0.235\phantom{)}$ & $0.663\phantom{)}$ & $0.058\phantom{)}$ &  &  \\
 & $(0.017)$ & $(0.118)$ & $(0.129)$ &  &  \\
$(\log z)^2$ &  & $-0.024\phantom{)}$ & $0.015\phantom{)}$ &  &  \\
 &  & $(0.007)$ & $(0.008)$ &  &  \\
$\log \log z$ &  &  &  & $1.947\phantom{)}$ & $1.898\phantom{)}$ \\
 &  &  &  & $(0.148)$ & $(0.209)$ \\ \hline
$\sigma^2_D$ & $4.73\phantom{0)}$ & $4.64\phantom{0)}$ & $4.41\phantom{0)}$ & $4.67\phantom{0)}$ & $\phantom{0)}$ \\
$\sigma^2_P$ & $6.42\phantom{0)}$ & $6.26\phantom{0)}$ & $5.69\phantom{0)}$ & $6.33\phantom{0)}$ & $\phantom{0)}$ \\
$\pi(1000)$ & $0.24\phantom{0)}$ & $0.32\phantom{0)}$ & $0.27\phantom{0)}$ & $0.28\phantom{0)}$ & $\phantom{0)}$ \\ \hline\end{tabular}
\end{center}
\end{table}
%
For this model, sandwich standard errors were reported based on Spady's 
estimation of the Hessian and outer-product matrix of the gradient.  Our
attempt to replicate these results  employed the \proglang{R} \pkg{sandwich}
package as for the growth regressions in Section~\ref{sec:growth}.
The results are again shown in our Table~\ref{tab:rk}. 
Comparing these results with the published estimates we find somewhat larger 
discrepancies than for the quasi-Poisson regressions, but the results are
remarkably consistent.  We would conjecture that this degree of agreement 
would be rare in most instances where independent software was used to do
non-linear maximum likelihood estimation.

\section{Challenges and Conclusions} \label{sec:challenges}

From an economic perspective, the real challenge of reproducible econometric
research lies in restructuring incentives to encourage better archiving and
distribution of the gory details of computationally oriented research.
Technical progress in software and computer networking have dramatically lowered
the cost of reproducibility, but without stronger incentives from journals and
research funding agencies, further progress can be expected to be slow.
The JAE is exemplary in this respect since it has strongly encouraged data/software
archiving as well as replication studies.  It would be excellent if other journals
followed this lead.  Authors ultimately need to be convinced that it is in their
interest to provide detailed protocols for the computational aspects of their work.
This may require a sea change in attitudes about acknowledgment and citation 
practices.

Further technical progress can be expected in all of the realms we have reviewed:
more convenient archiving and version control, better tools for literate programming,
improved algorithms and user interfaces for statistical computing are all under active
development.  More rapid  diffusion of this new technology is what is really needed.

Web-based ``electronic appendices'' are increasingly common and this too is a welcome
development. However, further pressure by the journals, a better understanding
of the corresponding required quality standards by the scientific community, as well as
simplified automatic access would be very valuable.  
We are still far away from the Claerbout Principle, but good models do exist.  
WaveLab of \cite{repro:Buckheit+Donoho:1995} is a relatively early example,
the concept of a \emph{data compendium} suggested by \cite{repro:Gentleman+TempleLang:2007}
is another. Within economics there has been some discussion and evaluation of the
archival policies of the \emph{American Economic Review} and \emph{Journal of Money, Credit and Banking}
\citep[see][]{repro:McCullough+Vinod:2003,repro:McCullough+McGeary+Harrison:2006},
but much more is needed.

\section{Computational Details} \label{sec:computational}

Our results were obtained using 
\proglang{R}~2.14.0---with the packages
\pkg{lmtest}~0.9-30, \pkg{sandwich}~2.2-9,
and \pkg{MASS}~7.3-16---and
were identical on various platforms including Debian GNU/Linux (with a
2.6.26 kernel) and Mac OS X, version~10.4.10. {\TeX} Live 2007 was used for the original typesetting.
The full sources for this document (including data, \proglang{R} code and {\LaTeX} sources)
are available from \url{http://www.econ.uiuc.edu/~roger/research/repro/}. Hence, readers
can do as we do, not as we say, and fully reproduce our analyses.

\bibliography{repro}

\end{document}

### 2013-05-06 paper-data-restore-model.Rnw
using sweave

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{acm_proc_article-sp}

%\usepackage[latin1]{inputenc}
\usepackage[utf8]{inputenc}

\usepackage{hyperref}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{listings}

\newcommand{\conceptual}[1]{
%\textcolor{RoyalBlue}{#1}
\texttt{#1}
}

% TODO: wir sollten noch die Terminologie klï¿½ren. wie bezeichnen wir 
%
% * das Prinzip, Werte zu referenzieren:  Bisher nannte ich es Reference Model
%
% * so ein Daten-Template, das IDs anstelle von Werten enthï¿½lt: data reference template? ddocks template?
%
%
% Solange das noch nicht so ganz gefunden ist, kï¿½nnen wir einfach diese Tokens verwenden:
%
\newcommand{\refmodel}{
%data reconstruction model
data restore model
}
\newcommand{\datatemplate}{
data template
}

\begin{document}
\SweaveOpts{concordance=TRUE}

\title{
%On the Reproducibility of Statistical Research
%On the Reproducibility in Computational Statistics
A Data Restore Model\\
for Reproducibility in Computational Statistics
}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

\numberofauthors{3} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Daniel Bahls\\
       \affaddr{Leibniz Information Centre for Economics (ZBW)}\\
       \affaddr{Neuer Jungfernstieg 21}\\
       \affaddr{20354 Hamburg, Germany}\\
       \email{d.bahls@zbw.eu}
% 2nd. author
\alignauthor
Benjamin Zapilko\\
       \affaddr{GESIS - Leibniz Institute for the Social Sciences}\\
       \affaddr{Unter Sachsenhausen 6-8}\\
       \affaddr{50667 Cologne, Germany}\\
       \email{benjamin.zapilko@gesis.org}
% 3rd. author
\alignauthor
Klaus Tochtermann\\
       \affaddr{Leibniz Information Centre for Economics (ZBW)}\\
       \affaddr{D\"usternbrooker Weg 120}\\
       \affaddr{24105 Kiel, Germany}\\
       \email{k.tochtermann@zbw.eu}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
\date{14 May 2012}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

\maketitle

\begin{abstract}

Researchers are more and more requested to publish their scientific data sets for purposes of transparency, re-use, and reproducibility.
Particularly in economics and the social sciences, researchers often use sensitive statistical data that underlies protection policies that inhibit distribution to third party archives.
In addition, a substantial quantity of data sets is combined from one or more external providers, which easily complicates curation-related activities. %---when working with data copies
These circumstances give us reason to pursue a data restore model based on fine-grained referencing that allows to trace data provenance to the original archive that is in charge of curation.
One goal is to enable data publication in difficult settings.
%BZ: "this research" koennte verwechselt werden mit der research, die reproduzierbar sein soll. wie waers mit "our goal is to enable ..."
Moreover, we show how the gaps between data citation and code integration can be closed as to eliminate all manual efforts of arranging code and data for reproduction attempts. 
On this basis, we develop the requirements for a data restore model and elaborate a generic design and its implication on an overall data management infrastructure.
We further present a proof of concept implementation on the basis of a real-world example from economics.
Eventually we close with the vision of a data and code ontology that carries empirical models from paper to a re-usable semantic level.

\end{abstract}

%todo: categories, keywords, terms setzen
% http://dl.acm.org/ccs_flat.cfm
% http://www.acm.org/about/class/2012

% A category with the (minimum) three required fields
\category{H.3.7}{Digital Libraries}{Miscellaneous}
%A category including the fourth, optional field follows...
\category{I.2.4}{Knowledge Representation Formalisms and Methods}{Semantic networks}

\terms{Design, Algorithms, Economics}

\keywords{Research Data Management, Semantic Digital Data Library, Linked Data, Statistics} % NOT required for Proceedings

<<preliminaries, echo=FALSE, results=hide>>=
#library("lattice")
#library("MASS")
#library("lmtest")
#library("sandwich")
library(ddocks)

options(prompt = "R> ", width=10, continue = "+  ", digits = 4, show.signif.stars = FALSE)
cleanup <- FALSE
@

\section{Introduction}

Repeatability of research results is a fundamental criterion in science. % experiments are conducted in different environments and by different persons as to validate whether the results depend on considered parameters only.
Experiments must lead to the same findings when conducted by other persons or in other environments, under the premise of keeping relevant and considered parameters constant.
%This is very often not the case in computational research 
% sciencemag article on "Reproducible research in computational science"
In computational research, this remains unverified in many cases due to various reasons.
\cite{peng2011reproducible}
%One of them is the plain fact that the data on which the research is built often remains unavailable to others which gives rise for a more organized data management practice. \cite{Wood1298248}
One of them is the plain fact that the supporting data often is unavailable to others which gives rise for a more organized and open data management practice. \cite{Wood1298248}
While a scientific publication is commonly understood as synonymous with a peer-reviewed article alone, program code and data take the role of nice-to-have additional material only that rarely find their way to the public.
%There are numerous examples of research publications that turned out inconsistent eventually, either because of mistakes,  
%There are numerous examples of inconsistent research  that illustrate how

Transparency and verification of results not only are gestures of good scientific practice but essential if the statements published in the name of science are to reflect profound and solid knowledge.
%BZ: Hier koennte noch der von dir angesprochene punkt zu journals und datenpublikationen hin. Mehr als 1-2 saetze brauchts hier aber nicht.
The problem might be less significant in technical domains where relevant theories and best practices would be validated eventually as soon as they are applied.
Decision makers in public sector and industry, however, often rely on surveys and studies which do not necessarily have to stand proof when made use of. 
The case of a research study on cultural integration of immigrants in the UK  
shows that inconsistencies occur in the social sciences as well.
The data was available to the community \cite{dataset_FNSEM} which allowed other researchers to investigate and after all heavily criticize their publication as its results could not be reproduced. \cite{arai2011fragile}
To underscore their critique, they published the source code of their own analyses online and in this way called for clarification. % and revealed their difficulties in replicating.
The original authors responded with an errata in which they acknowledged the faults.
At the same time, they presented a modified version of the statistical model they had applied and managed to produce the previous findings again while admitting that ``the results are now less clear-cut and [they lost] statistical significance''. \cite{bisin2011errata} 
%
% btw, the paper is not retracted. the authors did not highlight anywhere around the homepage that there was an issue. they simply list the errata as another publication and that's it.
% it has 58 citations as of today, and the errata is not cited once at all.
% ich will hier aber kein aggressives politikum starten, wir wollen ja auch niemanden abschrecken und schreien, dass wir hier aufräumen wollen, oder?
%

Data availability was the key for verification in the example above.
Yet, when it comes to analyses and the methods applied, textual reports typically lack many of the algorithmic details needed for reproduction or re-use after all.
Altogether, the media type itself seems rather unsuitable for this purpose, which is why there are many voices in the scientific community promoting the idea of program code to be included as substantial artefacts along with the article.
In this sense, some researchers even view source code as the main contribution.
This viewpoint is often referred to as the \textit{Claerbout principle}:
"An article about computational science in a scientific publication is not the scholarship itself, it is merely advertising of  the scholarship. The actual scholarship is the complete software development environment and the complete set of instructions which generated the figures."
\cite{Buckheit1995b}
\cite{de2001reproducible}
In support of this view \cite{KoenkerZeileis2009} investigated reproducibility in econometrics and pointed out how software tools can be used to organize data, code, and documentation in this domain.
As part of a replication study they have also shown that important details are missing in the two example papers examined and filled in the blanks by publishing source code and data of the eventually successful replication attempts.

Both, data and code resources make a valuable contribution if findings are to be reproducible and re-usable in computational research.
And yet, there is more to be considered.
%A comprehensive review of the JMCB\footnote{Journal of Money, Credit, and Banking} data+code archive reveals that results cannot be reproduced even when these resources are given.  
A comprehensive review of the \textit{JMCB\footnote{Journal of Money, Credit, and Banking} data+code archive} reveals that provenance information for the data used is often incomplete and that submitted code does not reproduce the results either.
It is not uncommon that even the original authors have difficulties understanding the details of their own work only few months after publishing already.
Apparently, these issues are faced in many other areas of computational research too
% reproducibility is a problem in many domains
% bio informatics (case studies 100 articles..) mit leisch, sweave autor
\cite{hothorn2011case}
% empirical software engineering
\cite{gonzalez2012reproducibility}
% signal processing
\cite{vandewalle2009reproducible}
% ecology
\cite{cassey2006reproducibility}
% "replicating empirical results is an underemphasized, yet essential element of scientific progress."
% Report on the American Economic Review Data Availability Compliance Project
% http://www.aeaweb.org/aer/2011_Data_Compliance_Report.pdf
which can be understood as a problem of insufficient documentation.

With the increasing use of statistics software \cite{RePEc_hum_wpaper_sfb649dp2011_020}
these problems emerge in the domain of empirical research in general where more and more statisticians promote the use of literate programming techniques  
% Executable Papers for the R Community: The R2 Platform for Reproducible Research
\cite{leisch2011executable}
% literate statistical practice (reproducibility and documentation)
\cite{rossini2003literate}
as to combine report text and coding in single executable scripts so that basic documentation of the analyses is already given naturally.
In both of the examples above, the authors made use of the literate programming tool Sweave \cite{lmucs_papers_Leisch_2002} 
so that the report can be generated again from data processing to generating the report document, under the premise that the data set is given and the system environment allows for correct execution.

However, empirical research is often based on sensitive data that cannot be shared easily.
%Such data sets can be transformed into anonymized public-use files in general which needs verification in terms of cell sizes and de-anonymization and requires additional efforts, often on the researchers' side.
A general solution to this problem is to transform such data sets into anonymized public-use files that can be distributed without infringements of data protection.
But doing so requires careful verification in terms of cell sizes and de-anonymization and requires additional efforts, often on the researchers' side.
Moreover, researchers hesitate to publish their own data if incentives are not clear, especially when they had a lot of work producing it.
In addition, data is often obtained from several sources and the mere citation of them does not give any details on the file formats used or way of composition as used for input for the computer program.
%BZ: hier kommst du wieder zurueck zu den daten (wenn auch mit anderem schwerpunkt), nachdem das thema oben eigentlich abgehakt wirkte und du zum thema program code und danach zu provenance gegangen bist. sollte dieser abschnitt ueber sensitive data nicht zu dem data abschnitt weiter oben? gehoert ja irgendwie schon zusammen, oder?

%These circumstances motivate our efforts to elaborate on a holistic solution for the statistical domain that enables the exact reconstruction of individual research data sets as were used by the researchers.
In this paper we aim to overcome these challenges for the domain of statistical data and elaborate on a holistic solution that enables the exact reconstruction of individual research data sets as were used by the researchers.
We pursue the idea of a \emph{data restore model} that is based on fine-grained data referencing as described in earlier works \cite{BahlsTochtermann_2012}
where further background and motivation is explained more thoroughly.
As this referencing technique can be viewed as means for precise citation, the linkup to provenance and further documentation on the particular data version is given which again contributes to clarity.
The idea is to equip researchers with means that make full transparency possible even in difficult settings of protected and distributed data sources, so that reconstruction remains a question of access permission only.
%We aim to design this referencing technique in a way that allows for seamless integration with preferred research tools and source code, so that no further manual work is required for setting up data and code resources during replication attempts.
%The restore model is designed to seamlessly integrate with preferred research tools and source code, so that no further manual work is required for setting up data and code resources during replication attempts.
The restore model is designed to bridge the gaps between data citation and program code and to seamlessly integrate with preferred research tools, so that no further manual work is required in setting up data and code resources for replication attempts.
Eventually, the goal is to provide researchers with the ability to close all missing links from data sources and processing to figure plotting and report generation as to create trust in their findings which from another perspective could be an incentive to do so.
On the basis of this data restore model, we elaborate the requirements it carries on the entire information infrastructure between researchers, archives, and libraries.

%In this paper we aim to overcome these challenges for the domain of statistical data and elaborate on a holistic solution for an infrastructure that enables reconstruction of individual research data sets. 
%This infrastructure is basically realized by an underlying data reconstruction model, which builds the bridge between the researchers' preferred software tool and the research data. 
%The approach follows the idea of a data reconstruction model by which we develop the requirements it carries on an infrastructure.
%We present an implementation of the model that tries to seamlessly integrate with any software tool preferred by a researcher.
%By utilizing the model generic ids are assigned to all data values regardless of their original format and their physical location.
%This allows for reconstructing and referencing the data set out of the researchers programming tool.

The remainder of the paper is structured as follows. 
In section 2 we present a case study with real world examples, where literate programming with research data is applied. 
Based on these examples we propose a generic data reconstruction model in section 3. 
We implement and validate our model in section 4. 
In section 5 we discuss related work relevant to our approach. 
Finally, we conclude and present future work in section 6.
%BZ: veraltet. muss man am schluss aktualisieren, wenn das paper steht

\section{Related Work}

%argumentation: our tpoic is related to several fields and tries to connect lose ends. topics: general reproducibility, data access and privacy, data referencing/citation

%------------------------------------
%topic general reproducibility

% stichpunkte, was hier noch rein kann
% da muss man mal sehen, ob das wirklich alles rein muss
The problem of reproducibility in computational research is addressed by many.
If re-execution of experiments is the goal, it is necessary to capture information on the entire system environment used as to make sure respective processes can be performed under the same parameters.
%
% preservation of digital objects and processes (problems with system environments)
%
%
% Virtual Machines: eine Virtuelle Maschine wird gespeichert (ungefähr die Ecke Andreas Rauber, Klaus Rechert)
% preservation of processes and digital objects
\cite{DBLP_conf_ercimdl_Rauber12}
\cite{Rechert_2010_EBS_1816123_1816182}
%\cite{Guttenbrunner_2012_MFE_2180868_2180876}
%This alone is already a very hard problem, because it must include hardware and software dependencies to very detail.
This alone is already a very hard problem, because the chain of dependencies in terms of hardware and software has to be captured and restored to the very detail.
In this regard, \cite{Schwab_2000_MSC_369545_369555} illustrates how the problem of software dependencies for computational research could be solved with the help of build scripts such as make files.
Nevertheless, the achievements in this area seem promising, and the opportunities for application such as  
%
%
% Virtual Research Environments / E-Science, Grid Computing
%
% collage is an executable paper environment to re-run analyses (grid-based)
\cite{Nowakowski2011608} % waren die 2. gewinner der exec paper challenge 11
%
% scientific workflows passen hier auch
are great. % todo: echt beknackte formulierung
%BZ: wie waers mit:
%Nevertheless, the achievements in this area and the opportunities for application such as \cite{Nowakowski2011608} seem promising.
As the goal of our research is to find means for precise reconstruction of data sets and direct integration with program code, these approaches form a good complement. % todo: formulierung   
%BZ: As we aim to find means for precise reconstruction of data sets and direct integration with program code, the approaches complement each other.

Particularly for the domain of empirical research, we found the literate programming approach very appealing, as it seems light-weight and suitable for less complex scenarios that can be handled with a single script.
It interweaves source code and text for the report, so that all processing steps are already quite well-documented. 
%Approaches like \cite{lmucs_papers_Leisch_2002} prove that the reproduction of research results is technically feasible, but there are still challenges that hinder this process: 
%Literate programming approaches like Sweave \cite{lmucs_papers_Leisch_2002} or Knitr\footnote{\url{http://yihui.name/knitr/}} are suitable for the reproduction of research results. 
Tools like Sweave \cite{lmucs_papers_Leisch_2002} or Knitr\footnote{\url{http://yihui.name/knitr/}} are already accepted as helpful means to reproduce analyses and document its steps
% Executable Papers for the R Community: The R2 Platform for Reproducible Research
\cite{leisch2011executable}
% quite a number of recent publications make use of them.\footnote{The list of Sweave citations in scientific publications of many areas is increasing continuously.}
% the author's initial motivation was to be able to re-run his own analyses again
%
% literate statistical practice (reproducibility and documentation)
\cite{rossini2003literate}
and therefore suitable for the reproduction of research results.

%
% weitere evidenzen fuer die beliebtheit als reproduzierbarkeits-tool sozusagen. brauchen wir das?
%
% literate programming for transparency
% Encoding Transparency: Literate Programming and Test Generation for Scientific Function Libraries
%\cite{flood2012encoding}
%
% A Multi-Language Computing Environment for Literate Programming and Reproducible Research
%\cite{Schulte2012}
%
% Literate Spreadsheets (nicht gelesen ;)
%\cite{dinmore2012design}

% noch mehr kram, "Reproducible computational experiments using SCons"
% cite{fomel2007reproducible}

Source code for computational research does not need to be executable in order to make a valuable resource for transparency,
as it is a detailed documentation of the processing already and a valuable resource as such, which is part of the argumentation in 
% all these resources should have an identifier, as the title suggests
% they won the executable paper challenge, although it is not executable. it says quite something
%
% verifiable computational research
%
\cite{gavish2011universal}.
The Verifiable Computational Research community aims to collect data, code and intermediate results of research processes as a trace to make the findings verifiable.
Their automatic capturing process is integrated with the researchers' workflow, and therefore successful as well as unsuccessful attempts can be reviewed eventually. 
%However, the VCR community argues that the source code does not have to be executable, source code as a detailed documentation of the processing suffices, and other intermediate results should be captured, so that the research is tracable and verifiable.
We share their view in a pragmatic sense. The code does not need to be executable, if it is, the better. % todo: formulierung 
%BZ: Ich wurde den letzten teil ab "if it is" einfach streichen...
Either way, source code should be picked up as part of the publication material whenever possible.
%
% source code is important 
%
% * for replicability not only data itself, but also program code, e.g. of mathematical calculations, is necessary
%
% science code manifesto
\cite{barnes2012science}\footnote{\url{http://sciencecodemanifesto.org/}}

%------------------------------------
%topic data access and privacy

%wenn das als related work durchgehen soll brauchen wir hier noch mehr referenzen...

%One of the main problems in empirical research, however, is the fact that a great deal of the data used is sensitive and hence protected.
%Other data sets are used commercially and usage policies prohibit its further distribution
One of the main problems in empirical research, however, is that a large amount of the data used is sensitive and hence protected, or underlies certain usage rights due to commercial use. % REF da gibts bestimmt irgendwas vom RatSWD oder so...
Therefore the problem of data availability cannot be solved as the legal situation simply prohibits the further distribution to other parties.
% open data
But recently, as a result of the Open Data movement a large degree of data sets has been made publicly available by governments, statistics offices and research organizations. This data can be used for research, but is not always citable or referencible in a way it is required for scientific publications like using persistent identifiers. The publictaion of such data as Linked Open Data \footnote{http://linkeddata.org/} allows a precise identification and recerence of single entities such as data sets or particular indicators. Furthermore, by using the Linked Science Ontology \cite{kauppinen_et_al_linked_science_2011} abstract research processes can be modelled and represented. But, this ontology does not address the issues of data access and availability.

%

%

% panton principles
\cite{murray2010panton}

% CKAN muss auch noch irgendwo rein
% das ist doch aber immer open data.
% wir können es nutzen um die templates zu publizieren, ansonsten kommt es wieder in die rubrik: geht nicht, datenschutz

%
% LOD
%
%  * LOD for scientists.
%
%\cite{Halb_LDOW08}
%
% * Linked Data: dazu sollten wir etwas mehr sagen, aber das hast du ja oben schon etwas ausgeführt
%   knüpft an Metadaten an, Ontologien wie Data Cube etc, semantic Web
%
% da passt auch die linked science ontology
%\cite{kauppinen_et_al_linked_science_2011}

%--- der rdf teil hier kann eigentlich oben mit in die related work oder?

%Represented in RDF, using the RDF Data Cube Vocabulary (QB) for example, documentation on the variables could be interweaved within and machines could make meaning of it too, if given according to established standards like SDMX, DDI or alternative application-oriented ontologies.
%%There are other initiatives that like to see this data in this structure too
%It would integrate with other open government data.
%There is interest from the linked data community as well. \cite{Halb_LDOW08}
%Eventually, RDF provides strong means for referencing as it builds on URIs\footnote{Or IRIs} which allows for decentrally organized ID assignment at no costs.

% why linked data is not enough for scientists, also reproducibility etc.
%\cite{bechhofer2010linked}

%
% other related work:
%
% * data provenance: können wir anstoßen und sagen, dass wir ja egal was referenzieren
% * genauso auch die Diskussionen um Metadaten: DDI, SDMX, alles in eine Box und sagen, dass wir es hier nicht brauchen
%

% data cite
%\cite{Citation2011}
% SDMX, DDI

%BZ: provenance und metadaten würde ich hier glaub ich rauslassen, das spannt den bogen sonst zu weit. habe auch lod nur kurz mitaufgenommen.

%------------------------------------
%topic data referencing / citation

%Data citations typically point to entire data sets as a whole.

%BZ: hier könnte eigentlich ein bißchen allgemeinsa zu data referencing / citation hin... das fängt oben schon ein wenig bei open data und lod an. dann könnte man sich im folgenden kapitel nämlich ein wenig kürzer fassen und besser auf das model fokussieren.

\section{The Data Restore Model}\label{sec:datareferencing}

In this section, we first develop the requirements for a data restore model.
%Subsequently, we elaborate a generic design and outline the big picture of the overall data mangement infrastructure and workflows.
Subsequently, we elaborate a generic design and discuss its implication on an overall data mangement infrastructure and the workflows involved.

\subsection{Requirements}

Whichever data management solution is applied eventually, it should fit to the researchers' workflows and practices in order to be effective and accepted hence.
\cite{Feijen_2011} 
In the ideal case, researchers can use their preferred tools with no further hassle preparing anything at the time of data submission.
%  (assisting tools to reduce efforts)
%
%On the side of reuse, there should not be any unnecessary efforts in re-using these resources either, meaning that no puzzling and manual tinkering in organizing data and code to make it run.
On the side of reuse, there should not be any efforts needed in composing and organizing these resources either, meaning that no puzzling and manual tinkering in organizing data and code to make it run is required.
Location and structure of data files on the local file system should be inherently specified, so that it integrates perfectly with the code.
\footnote{In this regard, software and system specifications are as important. This kind of documentation can be organized independent of the issues addressed in this paper, we therefore do not further address it here.}

There are plenty of prevailing file formats, ranging from CSV, XML, Excel, entire RDBMS dumps to specific software-dependent and even unknown individual ``raw" types of formats.
McCullough et al. recommend the use of ASCII encoding as to support re-usability \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337}, we therefore want to exclude binary formats from our discussion.
As to be less restrictive and support more contemporary encodings such as UTF-8, we want to pursue the idea of using self-describing documents with information on its own encoding contained.
\footnote{Otherwise, data would need to be converted back and forth which causes trouble whenever a character has no representation in the ASCII standard}
However, a format-independent solution should be found whereas most importantly, any restored data set must equal the original as to perfectly integrate with the code in the end. 
% which gives us a clear (requirement for successful) test case
Thus, we note that data citation by persistent identifier (as is the current practice in data publishing) is not sufficient in this regard, since these identifiers point to data sets on a more abstract level only and do not distinguish or specify a particular format.
\footnote{Most data agencies offer download in various formats while the content is filed under the same persistent identifier.}

Further, the data restore model should be applicable for protected data as well.
This might sound surprising to the reader as sensitive data cannot be shared per se and a restore concept appears redundant.
Nevertheless, we want to make the point that the spectrum from open to protected data is gradual, since some data is access-restricted only and can be obtained under certain policies and regulations.
Whether or not a data set can be handed out should be a question on legal level only, the restore technique itself, however, should cater for these cases as to give means for repeatability in every scenario.
% später nochmal aufgreifen: meta data von protected data sharen. man kann wenigstens herausfinden, dass es diese daten gibt

Either way, pointers to the provenance of the data used should be available that give documentation and further background such as contact information to creators and curators for any version being used.
While new persistent identifiers are commonly given for major data updates and releases only, those versions with minor bug-fixes cannot be referred to with precision so far.\footnote{Surely, an author can give a note on the time of data download or acquisition in the paper. This is rarely practiced, and data reconstruction in this case again involves vagueness and human efforts.}
%A common practice is to assign new persistent identifiers to major data updates while those versions with minor bug-fixes cannot be referred to so far.\footnote{Surely, an author can give a note on the time of data download or acquisition in the paper. This is rarely practiced, and data reconstruction in this case again involves vagueness and human efforts.}
%It is also to consider that new persistent identifiers are given for major data updates only while those versions with minor bug-fixes cannot be referred to with precision so far.\footnote{Surely, an author can give a note on the time of data download or acquisition in the paper. This is rarely practiced, and data reconstruction in this case again involves vagueness and human efforts.}
The use of versioning systems could help identify individual releases at detailed level \cite{KoenkerZeileis2009} which should integrate seamlessly with the data restore model.
Eventually, references to service points providing this information should be given along with the submission material.
This could further be realized with Web Services that accept requests for information or data content on a technically fully supported level, so that up-to-date user assistance can be provided directly within the tools used in a research environment.
Depending on regulations, the actual data content identified by a given model instance could be transferred on request, possibly using authentication mechanisms, so that the data set can be restored locally using the original data sources.
\footnote{Authentication can be realized using SSL for example. If transfer via internet is considered too insecure, the API can be used to collect data requests as such and to initiate the procedure according to the respective policies.}

In contrast to the copy-based approach, the referencing technique allows data and documentation to be maintained at one responsible archive only, so that no duplicate efforts have to be made regarding data curation which is particularly relevant in cases where several external data sources were used.
This improves quality of service as all enquiries about the data can be answered with the highest possible level of expertise available and gives a clear view on responsibilities in addition.

As a result from this discussion we formulate the following requirements for the data restore model:

\begin{enumerate}\label{reqlist}

%  \item First and foremost, data sets should be reconstructable to the fullest, so that experiments can be repeated with the exact same data.
%  \item \textbf{Equality:} Data sets should be reconstructable to the fullest, so that experiments can be repeated with the exact same data.
  \item \textbf{Equality:} A data file restored by the model must equal its original on character level\footnote{Depending on character encoding and operating system, the files may differ on bitstream level while the textual content is equal.}, so that it integrates perfectly with the software and program code used by the researcher.

%  \item The technique must be applicable to protected or closed data as well.
  \item \textbf{Protection:} The model must be applicable for protected data (closed data) as well, in strict accordance with secured protection.

  \item \textbf{Curation:} The technique makes possible that data could be maintained and curated at those archives only that are responsible of maintenance with the required level of expertise.

  % das ist vielleicht etwas hart formuliert
%  \item Researchers are not disturbed by the introduction of technical hinderances such as new data formats or other obligations.
  \item \textbf{Usability:} The method should integrate seamlessly with the researchers' workflows and suit to their preferred data formats. 
		% It should be independent of any particular data format. % kann man das so reinschreiben, klingt jedenfalls komisch

  % auch soll es möglichst einfach auf neue unbekannte Datenformate anwendbar sein, mitunter um zukunftsfähig zu sein
  \item \textbf{Formats:} The method should be sustainable and applicable to alternative, unknown or future data formats.

  % verwendung von ids macht das ganze glasklar. kann man als argument gegenueber query-approach anbringen, wenn wir wollen 
%  \item It should be simple and transparent, so that problems can be traced easily. % mit hilfe der direkten pointer ist alles gestochen scharf und einfach zu debuggen (?)
  \item \textbf{Tracability:} The technique itself must be transparent, so that problems can be traced and questions can be answered efficiently.
 %wenn du mit traceable meinst, dass zb berechnungen auf den daten nachvollziehbar sein sollen, dann wuerde ich hier auf jeden fall die ids nennen. transparent ist hier dann als begriff zu allgemein.

% zu weich, was heißt das?  
%  \item overall, it has to meet the constraints and requirements from all parties involved in the overall data management infrastructure, such as archives, researchers and libraries. % was vergessen? und not openly available data = closed data?

  \item \textbf{API:} The data restore model should include pointers to Web Services that provide documentation, contact information or, if applicable, the actual data content upon request.

\end{enumerate}

%------------model--------
\subsection{Designing the model}
%\subsection{ID-per-Value Approach}

The requirements can be realized in several ways.
However, format-independent reconstruction of individual data files requires a mechanism that renders data content into any unknown or unspecified structure.
The referencing technique implies that data slices of larger data sets can be identified and selected precisely, from global scope.
Pointers to respective repositories in combination with SQL, SPARQL or other kinds of queries could be a solution.
A challenge in this approach is that such queries produce result sets in the shape of tables which yet need to be transformed to this unspecified format.
Whereas this idea is certainly worth exploring, we want to pursue another approach that fits better into the Semantic Web scheme and recent developments in the Linked Open Data movement.

Currently, more and more data sets are released as Open Government data and statistical offices have already explored the use of semantic technologies as well \cite{Halb_LDOW08}.
%BZ: Hier suche ich noch ein, zwei Referenzen raus, die zusammen mit Halb_LDOW08 aufgezaehlt werden koennen.
These technologies break down the separation between black box data bundles and attached metadata as it interweaves data with its documentation.
%BZ: These technologies break not only down the separation between black box data bundles (the raw data) and attached metadata as it interweaves data with its documentation.
%BZ: They also enable the publication, sharing and interlinking of data.
The data content is more accessible and interpretable to machines which sets the stage for a variety of new applications, e.g. for data retrieval. \cite{BahlsTochtermann_2012}
%BZ: das wesentliche fuer unseren approach ist hier eigentlich gesagt, alles andere wuerde wohl eher in die related work passen, oder? wuerde diesen absatz mit dem folgenden zusammenfassen.

Our approach is based on the concept of assigning individual IDs to every single value of a data set.
%BZ: Das ist notwendig, da man nicht das dataset als ganzen restoren moechte, sondern ja gerade die einzelnen data values benoetigt, um ggf. berechnungen nachvollziehen zu koennen. Bekomm das jetzt grad nicht griffig formuliert...
The number of IDs needed is certainly high and questions of scalability might rise immediately.
To give an estimation on the technical implications, the number of additional triples needed is twice the number of data values, or the same number only if linking with other semantic resources is not made use of.
Maintenance could be supported with tools and there should not be any additional efforts, provided that the data is maintained in a structured representation already.
%BZ: what tools?
However, we would like to have a discussion on this naïve approach and leave the investigation of scalability for now.

As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set could be referenced from any context (\textit{Tracability}).
The requirement item \textit{Curation} is addressed as well, because data can be referenced with precision and maintenance of every data can done in one place. 
Since the use of RDF allows us to add an additional link between ID and value, we can separate the two and disclose all information about the data in very detail in machine-understandable form, and yet protect the values (\textit{Protection}).
Furthermore, when every value can be referred to from anywhere, a simple technique can be applied to enable format-independent reconstruction of data files.
We therefore introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values which solves the requirements of \textit{Equality}.
The data template contains a header section in which we can specify encoding and tokens that are used to identify IDs within the template.
It enables the format-independent solution (\textit{Formats}) as this flexibility allows us to avoid syntactical clashes with the particular data format used.
In addition, we also declare Web Service addresses for every namespace used (\textit{API}) under the pragmatic assumption that every archive provides one such Web Service for data enquiries only.
At this point, we do not want to elaborate on the Web Service interface, however, we have used SPARQL endpoints in the proof of concept implementation presented in Section \ref{sec:impl}.

\subsection{Workflows}

If implemented in big scale as a holistic information infrastructure for statistical research data management, the presented approach has further implications on workflows and gives ideas for a big picture.
Data archives could offer the download of data templates along with the actual data, so that researchers have the precise references right from the beginning of their research.
Alternatively, researchers could pull the data on the basis of such data template from within their research tool, provided that a respective plugin is available.
The advantage in the latter case is that this procedure could be embedded within their program code, so that no further modifications of data and code need to be done at the time of publishing.
The researcher simply submits code and data templates used, and other researchers can rerun the same analysis without any manual arrangements or collection of resources, simply by executing the code.

In case the researcher used own data, the data would have to be submitted to the dedicated archive first where the respective data template and its IDs involved are generated first.
The archive returns the template to the researcher who can then publish it in the same way as described above.
This procedure must be supported with tools for mapping and generation of IDs.
Furthermore, these tools should provide for further semantic annotation of the content (using the RDF Data Cube Vocabulary for example) so that documentation about the data is available as well.

Figure \ref{fig:modelprocessing} outlines the overall process and suggests roles and responsibilities for archives and digital libraries.
The data itself would be maintained at the archives where Web Services provide access to up-to-date information and, if the policies allow, to its contents.
%Data templates and source code resources behind a scientific article could be maintained at the library

\begin{figure}[htb]
\centering
\includegraphics[width=.48\textwidth]{img/fig-model-processing.png}
\caption{
Using the \refmodel
}
\label{fig:modelprocessing}
\end{figure}

%BZ: die ersten beiden absaetze klingen sehr hypothetisch und zeigen eher verschiedene szenarien einer nutzung auf, wenn ich das richtig verstehe. ich wuerde in dem abschnitt glaub ich nur die grafik zeigen und anhand dieser dann einen konkreten ablauf darstellen (mit teilen aus den ersten beiden abschnitten). klar ist das dann in gewisser hinsicht auch hypothetisch, aber wenn die grafik zuerst kommt, viel anschaulicher.

%\section{A case study - Two real world examples}
%\section{An example from the economics}\label{sec:impl}
\section{Proof of Concept}\label{sec:impl}

%We have implemented
In this section, we present a prototype implementation of the data restore model and a plugin called \textit{ddocks}\footnote{\textit{ddocks} is short for data docks which is meant to reflect the ``transshipment" character of the interface between data retrieval and processing.} 
for the computing environment R.
We begin with the creation of data templates and IDs using two data sets from the sources of our example paper \cite{KoenkerZeileis2009}.
The plugin implementation is explained subsequently, and we close this section with a validation that tests reconstructed and original data files for equality.

For further details, the reader is invited to take a look at the source code we published on GitHub.
Respective links are given in the subsections
where we explain data template creation, the \textit{ddocks} plugin and validation steps which are part of the Sweave sources for this article.

\subsection{Generating the Data Templates}

We used a simple script that scans any given input text file for contained values using a regular expression which was specially designed for the data of our example.
For every identified value a static counter variable increments by one which is used to generate the value ID.
We could have used the plain counter value as such, but we decided to hash it only to highlight that these IDs need no meaningful structure at all.
In combination with the examplary archive namespace \textit{http://demoarchive.demo/data/} we obtain full URIs for all values, making them referenceable worldwide from any context.
To obtain the \textit{body} of a data template, all values are replaced with their abbreviated URIs and surrounded with angle brackets for parsing purposes.
A \textit{header} section is prepended to this part which we start with an XML-like notation that declares character encoding and \textit{ddocks} version as to give basic information about how to interpret the contents of this file.
The \textit{ddocks-header} tag opens the actual header section in which the namespaces used are listed with their associated URI prefix and repository API for which the prototype expects the address of a SPARQL endpoint.
It further declares the parse tokens used to identify IDs in the template body, so that clashes with format-specific syntax can be avoided.
Eventually, we append the body section right after the newline character behind the close tag of the header section to complete the data template.
Figure \ref{fig:datatemplate} shows an excerpt from one of the two data templates we have created.

\begin{figure*}[htb]
\centering
\includegraphics[width=.95\textwidth]{img/datatemplate.png}
\caption{
Excerpt from one of the data templates
}
\label{fig:datatemplate}
\end{figure*}

%
%
%\begin{quote}

\url{https://github.com/dbahls/ddocks-utils.git}
%\end{quote}

%For a more detailed view on the implementation, please take a look at the source code.\footnote{\url{https://github.com/dbahls/ddocks} \url{}}

\subsection{The R-Plugin}

The signature of \textit{ddocks\_restore}, the only function offered by the plugin, expects two parameters.
The first parameter is a connection object from which the data template will be read.
The second parameter specifies the target file the restored content will be saved to.

First, \textit{ddocks} interprets the information given in the header.
Using the declared open and close tokens, it collects all IDs occurring in the template body and organizes them by namespace which is the key to identify the respective data repository.
Iterating over these namespaces, the values are queried from the associated SPARQL endpoint\footnote{This is done in bundles so as not to query every single value separately. Then again, as to avoid oversized queries, these bundles are limited to a number of 30 IDs per query.}
and replace the IDs in the template body which is then written to the target file.

We run this procedure for the two data templates we have generated earlier
<<results=hide>>=
ddocks_restore(
  file("./data-templates/data.dj.ddocks"), 
  targetfile="data.dj")
@

<<results=hide>>= 
ddocks_restore(
  file("./data-templates/rk.raw.ddocks"), 
  targetfile="rk.raw")
@
and reproduce the data files on the local file system.
\textit{ddocks} takes these files as a cache and skips when the restore command is run again, so that the data does not have to be reconstructed every time the code is executed.
Researchers could embed these calls within their source code to load the data by template at the beginning of their analyses and leave it there without being bothered.
%This allows us to comfortably use the code within the source code of a research project from the begin
%
%
%\begin{quote}

\url{https://github.com/dbahls/ddocks.git}
%\end{quote}

\subsection{Validation}

For a basic proof of concept we have to show that the restored files equal their originals on character level.
If the two expressions
<<>>=
all.equal(
  readLines("data.dj"), 
  readLines("koenker-zeileis-09/data.dj"))

all.equal(
  readLines("rk.raw"), 
  readLines("koenker-zeileis-09/rk.raw")) 
@
evaluate to TRUE we have successfully validated the approach in terms of functionality.

We also inserted the two restore calls above in the source code of our example paper (\cite{koenkerzeileis09}) right before the load-data calls respectively, and running Sweave on it produced a report identical to the published article.\footnote{using R x64 2.15.3 on Windows}
%
%
%\begin{quote}

\url{https://github.com/dbahls/paper-restore-model.git}
%\end{quote}

\section{Vision, Summary, and Outlook}

The ideas presented in this work could be used to offer data and code behind statistical research publications to students and researchers.
We have demonstrated how detailed referencing can be used in a data restore model that 
 * overcome problem of data protection

enables the

The material could be used in tutorials to teach computational research in practice, serve as re-usable research resources, or could be used within review-processes.

or as part of review-pr

We have overcome some of the obstacles (data protection, distributed sources, tracability, source code integration (vs. citation) ) and shown how a "one-click solution" to reproduce results can be implemented in which the gaps between data citation and code integration are closed.

Tutorials could be 
The material could help learn and understand research in practice

%BZ: die zusammenlegung finde ich sehr sinnvoll

%
% Metadaten, DDI, SDMX, LOD etc. können wir im Prinzip als Technologien zur Dokumentation und zur Unterstützung von Data Retrieval betrachten.
% Wir suchen aber ja nicht nach Daten sondern haben konkrete Pointer. Wsa anderes wird in der Reproduktionsfrage auch nicht benötigt.
% Auskehren, dass unsere Technik sich prima mit Sem Web Technologien integriert. 
% Die Diskussion um geeignete Metadaten und Ontologien sollte vielleicht unabhängig von unserer Problemstellung geführt werden?
%
% Andererseits geht es auch um die Integration des Modells mit der Literatur. 
% Irgendwie wollen wir ja schon mit reinbringen, dass Template und Source Code dann auch mit einem Artikel assoziiert werden sollte.
% Dafür brauchen wir eine Ontologie.

% Das hängt auch mit der möglichen Anwendung zusammen, also
%  * Zielgruppe Studenten, Doktoranden, Wissenschafter, als Lehrmaterial allgemein oder für effizientere Forschung
%  * Review-Prozesse
%
% So eine Ontologie könnte es erlauben Daten und Code so zu verweben, dass man gut anbieten kann anstelle der geschützten Daten einen offenen Zwischendatensatz zu verwenden, mit dem weitergerechnet werden kann oder so....
%

%
% Input für eine Data and Code Ontologie
%

% Future work will be on that ontology too (name "data and code ontology"?)
% app scenario, for students, researchers, reviewers
% responsibility of libraries to catalog these items (restore model instances, instances of that ontology)
% archives should pick up the resources and apply restore model, if not ours then another one that is as effective

% dabei berücksichtigen wir bereits aufgestellte anforderungen

%
% Anforderungen aus Economics und Social Sciences zur Verbesserung der Reproduzierbarkeit
%
% Code ist immer erwähnt. und sie sagen, wie die readme datei aussehen muss usw.
% Diese ganzen elemente können in einer ontologie organisiert werden 
%

there are other recommendations in \cite{kleiber2012reproducible}   % replication files, code, technical information
that give requirements for an ontology that organizes the data and code resources.

%
% recommendations (liste von ressourcen, die gegeben sein sollten, damit reproduktion möglich ist)
%
\cite{KoenkerZeileis2009}
% mccullough, got replicability
\cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337}

% further guidelines
\cite{baiocchi2007reproducible}

% Gentleman: "important, if not essential, to integrate the computations and code used in data analyses,...."
\cite{gentleman2007statistical}

% habs nicht gelesen, geht aber direkt mit reproduzierbarkeit los. muessen mal sehen was die da an guidelines reinbringen, deckt sich bestimmt mit den anderen
% Reporting guidelines for simulation-based research in social sciences
\cite{rahmandad2012reporting}

%
% Zudem könnte man versuchen obige guidelines noch weiter ausbauen und folgendes mit aufnehmen
%
%  * Spezifikation von Systemumgebungen - Anknüpfung an Rauber, 
%  * Maven, RPM, dependency management
%  * Replikationsversuche (steht schon in den guides)
%

Scalability needs to be tested. If the idea is rolled out to the archives, would the servers with billions of IDs scale?
% db: ich bin da nicht wirklich skeptisch, aber die frage muss schon mal rein
% db: naja, das ist schon ganz schön heavy, was da an ID-Mengen entsteht. Wir sollten einfach einräumen, dass wir auf die Frage noch keine Antwort haben
%However the scalability question also applies for all the other RDF approaches in OpenGov, LOD etc, see billion triple challenge

We invite the reader to take a look at the sources for this publication and form her own opinion.\footnote{\url{https://github.com/dbahls/ddocks}}
% The final version of the paper will be using Sweave, for which the sources will be available as well. \footnote{\url{https://github.com/dbahls/paper-src}}
% db: die müllige revision history braucht man ja nicht in die welt schreien  ; )
% We used Sweave for this publication ourselves and invite you to make your own thoughts

\balancecolumns

\bibliographystyle{abbrv}
\bibliography{bibliography}

\end{document}

### 2013-05-06 paper-data-restore-model.tex
using sweave

--- 2013-05-06 paper-data-restore-model.Rnw
@@
 
+\usepackage{Sweave}
 \begin{document}
-\SweaveOpts{concordance=TRUE}
+\input{paper-data-restore-model-concordance}
 
@@
 \keywords{Research Data Management, Semantic Digital Data Library, Linked Data, Statistics} % NOT required for Proceedings
-
-<<preliminaries, echo=FALSE, results=hide>>=
-#library("lattice")
-#library("MASS")
-#library("lmtest")
-#library("sandwich")
-library(ddocks)
-
-options(prompt = "R> ", width=10, continue = "+  ", digits = 4, show.signif.stars = FALSE)
-cleanup <- FALSE
-@
 
@@
 We run this procedure for the two data templates we have generated earlier
-<<results=hide>>=
-ddocks_restore(
-  file("./data-templates/data.dj.ddocks"), 
-  targetfile="data.dj")
-@
+\begin{Schunk}
+\begin{Sinput}
+R> ddocks_restore(
++    file("./data-templates/data.dj.ddocks"), 
++    targetfile="data.dj")
+\end{Sinput}
+\end{Schunk}
 
-<<results=hide>>= 
-ddocks_restore(
-  file("./data-templates/rk.raw.ddocks"), 
-  targetfile="rk.raw")
-@
+\begin{Schunk}
+\begin{Sinput}
+R> ddocks_restore(
++    file("./data-templates/rk.raw.ddocks"), 
++    targetfile="rk.raw")
+\end{Sinput}
+\end{Schunk}
 and reproduce the data files on the local file system.
@@
 If the two expressions
-<<>>=
-all.equal(
-  readLines("data.dj"), 
-  readLines("koenker-zeileis-09/data.dj"))
-
-all.equal(
-  readLines("rk.raw"), 
-  readLines("koenker-zeileis-09/rk.raw")) 
-@
+\begin{Schunk}
+\begin{Sinput}
+R> all.equal(
++    readLines("data.dj"), 
++    readLines("koenker-zeileis-09/data.dj")) 
+\end{Sinput}
+\begin{Soutput}
+[1] TRUE
+\end{Soutput}
+\begin{Sinput}
+R> all.equal(
++    readLines("rk.raw"), 
++    readLines("koenker-zeileis-09/rk.raw")) 
+\end{Sinput}
+\begin{Soutput}
+[1] TRUE
+\end{Soutput}
+\end{Schunk}
 evaluate to TRUE we have successfully validated the approach in terms of functionality.

### 2013-05-06 paper-data-restore-model.Rnw
...

@@
 In addition, a substantial quantity of data sets is combined from one or more external providers, which easily complicates curation-related activities. %---when working with data copies
-These circumstances give us reason to pursue a data restore model based on fine-grained referencing that allows to trace data provenance to the original archive that is in charge of curation.
+These circumstances give us reason to pursue a data restore model based on fine-grained referencing that allows to trace data provenance to the original archive in charge of curation.
 One goal is to enable data publication in difficult settings.
@@
 Moreover, we show how the gaps between data citation and code integration can be closed as to eliminate all manual efforts of arranging code and data for reproduction attempts. 
-On this basis, we develop the requirements for a data restore model and elaborate a generic design and its implication on an overall data management infrastructure.
+On this basis, we develop the requirements for a data restore model and elaborate a generic design in view of an overall data management infrastructure.
 We further present a proof of concept implementation on the basis of a real-world example from economics.
-Eventually we close with the vision of a data and code ontology that carries empirical models from paper to a re-usable semantic level.
+Eventually we close with the vision of a data and code ontology that carries statistical models from paper to a re-usable semantic level.
 
@@
 
-\section{Vision, Summary, and Outlook}
+\section{Summary, Vision and Outlook}
+
+The ideas presented in this work can be used to offer data and code behind research publications, which increases transparency, re-use, and reproducibility.
+
+We have formulated requirements
+
+We have demonstrated how fine grained data referencing in our data restore model can overcome the problem of data protection.
+%Within the data template single data values can be addressed, even if the original data set is not publicly available.
+The introduced data templates provide precise pointers to the data used and can be disseminated freely as they do not disclose sensitive content.
+Since data files can be restored in the format used by the researcher, the template technique closes the gaps between data citation and program code integration.
+With the precise identification of values, provenance of the data used can be clarified comprehensively whereas responsibilities for maintenance and documentation of a data set are left to one archive respectively, so that harvesting techniques can be used efficiently to carry forward documentation, notes and updates about the data.
+
+%Researchers can make use of these information resources directly from within their research environment and run on updated data....
+
+We have overcome some of the obstacles (data protection, distributed sources, tracability, source code integration vs.
+citation) and shown how a "one-click solution" to reproduce results can be implemented in which the gaps between data citation and code integration are closed.
+
+Assigning URIs to single data values sets the stage for integration with RDF which has potential for valuable applications.
+A semantic model could offer researchers to run same analyses on updated data with no efforts on data retrieval and composition.
+While most statistical models today exist on paper only, the current work on reproducibility allows to rerun it again on the original or updated data, and the vision is to make them re-usable for other data as well.
+A research publication could remain more relevant for extended periods of time if a prognostic model on time series data for example can still be used on new records while the published article is no longer relevant to the community.
+Linked with thesauri, researchers can find these models by topic and re-use them in their own context while valuable user assistance can be given automatically if the semantic annotations specify clearly the kind of data it can process meaningfully.
+
+Moreover, these models could serve as input for the artificial intelligence community as they often describe abstract concepts like "living standards" or "justice" which are typically hard to model.
+Although these models are and should always be debatable, the semantic linkage to statistical indicators gives information on aspects that are relevant for understanding these concepts.
+Further, prognostic models could be combined in ensembles and trained with machine learning techniques, which might again serve the computational statistics community in their research.
+
+%
+% Vision
+%
+
+The material can be used in tutorials to teach computational research in practice, serve as re-usable research resources, or can be used within review-processes.
+
+%BZ: was genau meinst du mit the material? unseren code? restored data (references)?
+
+Considering a tight connection of data or data references with code and the eventually text of a research publication, the idea of a data and code ontology comes up.
+
+Such an ontology could be used to model the interconnection between data, code and research results in detail.
+Moreover, if also containing the data values of our data template, it could also be used as an intermediate data set for reproduce data computations without having access to the original data.
+
+We will investigate the design and development of a data and code ontology in our future work.
 
@@
 
-We have overcome some of the obstacles (data protection, distributed sources, tracability, source code integration (vs.
-citation) ) and shown how a "one-click solution" to reproduce results can be implemented in which the gaps between data citation and code integration are closed.
-
 Tutorials could be 
@@
 
-%BZ: die zusammenlegung finde ich sehr sinnvoll
+Scalability needs to be tested. If the idea is rolled out to the archives, would the servers with billions of IDs scale?
 
-%
-% Metadaten, DDI, SDMX, LOD etc. können wir im Prinzip als Technologien zur Dokumentation und zur Unterstützung von Data Retrieval betrachten.
-% Wir suchen aber ja nicht nach Daten sondern haben konkrete Pointer. Wsa anderes wird in der Reproduktionsfrage auch nicht benötigt.
-% Auskehren, dass unsere Technik sich prima mit Sem Web Technologien integriert. 
-% Die Diskussion um geeignete Metadaten und Ontologien sollte vielleicht unabhängig von unserer Problemstellung geführt werden?
-%
-% Andererseits geht es auch um die Integration des Modells mit der Literatur. 
-% Irgendwie wollen wir ja schon mit reinbringen, dass Template und Source Code dann auch mit einem Artikel assoziiert werden sollte.
-% Dafür brauchen wir eine Ontologie.
-
-% Das hängt auch mit der möglichen Anwendung zusammen, also
-%  * Zielgruppe Studenten, Doktoranden, Wissenschafter, als Lehrmaterial allgemein oder für effizientere Forschung
-%  * Review-Prozesse
 %
@@
 %
-
-% Future work will be on that ontology too (name "data and code ontology"?)
-% app scenario, for students, researchers, reviewers
-% responsibility of libraries to catalog these items (restore model instances, instances of that ontology)
-% archives should pick up the resources and apply restore model, if not ours then another one that is as effective
-
-% dabei berücksichtigen wir bereits aufgestellte anforderungen
 
@@
 
-% habs nicht gelesen, geht aber direkt mit reproduzierbarkeit los. muessen mal sehen was die da an guidelines reinbringen, deckt sich bestimmt mit den anderen
 % Reporting guidelines for simulation-based research in social sciences
@@
 
-Scalability needs to be tested. If the idea is rolled out to the archives, would the servers with billions of IDs scale?
-% db: ich bin da nicht wirklich skeptisch, aber die frage muss schon mal rein
-% db: naja, das ist schon ganz schön heavy, was da an ID-Mengen entsteht. Wir sollten einfach einräumen, dass wir auf die Frage noch keine Antwort haben
-%However the scalability question also applies for all the other RDF approaches in OpenGov, LOD etc, see billion triple challenge
-
-We invite the reader to take a look at the sources for this publication and form her own opinion.\footnote{\url{https://github.com/dbahls/ddocks}}
-% The final version of the paper will be using Sweave, for which the sources will be available as well. \footnote{\url{https://github.com/dbahls/paper-src}}
-% db: die müllige revision history braucht man ja nicht in die welt schreien  ; )
-% We used Sweave for this publication ourselves and invite you to make your own thoughts
-
 \balancecolumns

### 2013-05-06 paper-data-restore-model.Rnw
finishing!

@@
 %On the side of reuse, there should not be any unnecessary efforts in re-using these resources either, meaning that no puzzling and manual tinkering in organizing data and code to make it run.
-On the side of reuse, there should not be any efforts needed in composing and organizing these resources either, meaning that no puzzling and manual tinkering in organizing data and code to make it run is required.
+On the side of reuse, there should not be any efforts needed in composing and organizing these resources either, meaning that no puzzling and manual tinkering in organizing data and code to make it run is required \textit{(F partly)}.
+In the following, we want to develop the ideas of a referencing technique and take into account the recommendations \textit{(A-J)} fromulated by \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337} which we refer to using angle brackets.
 Location and structure of data files on the local file system should be inherently specified, so that it integrates perfectly with the code.
@@
 There are plenty of prevailing file formats, ranging from CSV, XML, Excel, entire RDBMS dumps to specific software-dependent and even unknown individual ``raw" types of formats.
-McCullough et al. recommend the use of ASCII encoding as to support re-usability \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337}, we therefore want to exclude binary formats from our discussion.
+McCullough et al. recommend the use of ASCII encoding \textit{(D)} as to support re-usability, we therefore want to exclude binary formats from our discussion.
 As to be less restrictive and support more contemporary encodings such as UTF-8, we want to pursue the idea of using self-describing documents with information on its own encoding contained.
 \footnote{Otherwise, data would need to be converted back and forth which causes trouble whenever a character has no representation in the ASCII standard}
-However, a format-independent solution should be found whereas most importantly, any restored data set must equal the original as to perfectly integrate with the code in the end. 
+However, a format-independent solution should be found whereas most importantly, any restored data set must equal the original as to perfectly integrate with the code in the end \textit{(E)}. 
 % which gives us a clear (requirement for successful) test case
@@
 
-Further, the data restore model should be applicable for protected data as well.
+Further, the data restore model should be applicable for protected data as well \textit{(supports G)}.
 This might sound surprising to the reader as sensitive data cannot be shared per se and a restore concept appears redundant.
@@
 Eventually, references to service points providing this information should be given along with the submission material.
-This could further be realized with Web Services that accept requests for information or data content on a technically fully supported level, so that up-to-date user assistance can be provided directly within the tools used in a research environment.
+This could further be realized with Web Services that accept requests for information or data content on a technically fully supported level, so that up-to-date user assistance can be provided directly within the tools used in a research environment \textit{(B)}.
 Depending on regulations, the actual data content identified by a given model instance could be transferred on request, possibly using authentication mechanisms, so that the data set can be restored locally using the original data sources.
@@
 
-As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set could be referenced from any context (\textit{Tracability}).
-The requirement item \textit{Curation} is addressed as well, because data can be referenced with precision and maintenance of every data can done in one place. 
-Since the use of RDF allows us to add an additional link between ID and value, we can separate the two and disclose all information about the data in very detail in machine-understandable form, and yet protect the values (\textit{Protection}).
-Furthermore, when every value can be referred to from anywhere, a simple technique can be applied to enable format-independent reconstruction of data files.
-We therefore introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values which solves the requirements of \textit{Equality}.
-The data template contains a header section in which we can specify encoding and tokens that are used to identify IDs within the template.
-It enables the format-independent solution (\textit{Formats}) as this flexibility allows us to avoid syntactical clashes with the particular data format used.
-In addition, we also declare Web Service addresses for every namespace used (\textit{API}) under the pragmatic assumption that every archive provides one such Web Service for data enquiries only.
-At this point, we do not want to elaborate on the Web Service interface, however, we have used SPARQL endpoints in the proof of concept implementation presented in Section \ref{sec:impl}.
+%As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set could be referenced from any context (\textit{Tracability}).
+%The requirement item \textit{Curation} is addressed as well, because data can be referenced with precision and maintenance of every data can done in one place. 
+%Since the use of RDF allows us to add an additional link between ID and value, we can separate the two and disclose all information about the data in very detail in machine-understandable form, and yet protect the values (\textit{Protection}).
+%Furthermore, when every value can be referred to from anywhere, a simple technique can be applied to enable format-independent reconstruction of data files.
+%We therefore introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values which solves the requirements of \textit{Equality}.
+%The data template contains a header section in which we can specify encoding and tokens that are used to identify IDs within the template.
+%It enables the format-independent solution (\textit{Formats}) as this flexibility allows us to avoid syntactical clashes with the particular data format used.
+%In addition, we also declare Web Service addresses for every namespace used (\textit{API}) under the pragmatic assumption that every archive provides one such Web Service for data enquiries only.
+%At this point, we do not want to elaborate on the Web Service interface, however, we have used SPARQL endpoints in the proof of concept implementation presented in Section \ref{sec:impl}.
 
@@
 for the computing environment R.
-We begin with the creation of data templates and IDs using two data sets from the sources of our example paper \cite{KoenkerZeileis2009}.
-The plugin implementation is explained subsequently, and we close this section with a validation that tests reconstructed and original data files for equality.
+Since literate programming already is a prominent approach to replicability in the computational statistics community \cite{leisch2011executable} (particularly supports recommendation \textit{(A)}) we want to demonstrate how the plugin can further strengthen this concept.
+%We begin with the creation of data templates and IDs using two data sets from the sources of our example paper \cite{KoenkerZeileis2009}.
+%The plugin implementation is explained subsequently, and we close this section with a validation that tests reconstructed and original data files for equality.
 
@@
 Respective links are given in the subsections
-where we explain data template creation, the \textit{ddocks} plugin and validation steps which are part of the Sweave sources for this article.
+where we introduce the concept of a data template using two data sets from the sources of our example paper \cite{KoenkerZeileis2009}. 
+The \textit{ddocks} plugin and a basic validation which are part of the Sweave sources for this article.
 
 \subsection{Generating the Data Templates}
+
+We introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values which solves the requirements of \textit{Equality}.
+The data template contains a header section in which we can specify encoding and tokens that are used to identify IDs within the template.
+In addition, we also declare Web Service addresses for every namespace used (which is only one for our example) under the pragmatic assumption that every archive provides one such Web Service for data enquiries only.
+%At this point, we do not want to elaborate on the Web Service interface, however, we have used SPARQL endpoints in the proof of concept implementation presented in Section \ref{sec:impl}.
 
@@
 \end{figure*}
+
+The script also produces an RDF document that contains the mapping from the URIs to their associated values.
+The simple ontology behind consists of one class only (\textit{DataReference}) and all its instances (value IDs) map to a literal (value) using the \textit{rdf:value} relation.
+Finally, we added these triples to an experimental triple store at GESIS that is presently open to the public.
+Please note that no further semantics were modelled as we wanted to demonstrate the minimal efforts necessary to follow the presented approach.
 
@@
 \textit{ddocks} takes these files as a cache and skips when the restore command is run again, so that the data does not have to be reconstructed every time the code is executed.
-Researchers could embed these calls within their source code to load the data by template at the beginning of their analyses and leave it there without being bothered.
+%Researchers could embed these calls within their source code to load the data by template at the beginning of their analyses and leave it there without being bothered.
+Researchers could embed these calls within their source code to load the data by template at the beginning of their analyses without being bothered, and no further modifications are needed at the time of publishing.
 %This allows us to comfortably use the code within the source code of a research project from the begin
@@
 
-For a basic proof of concept we have to show that the restored files equal their originals on character level.
+As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set could be referenced from any context (\textit{Tracability}).
+The requirement item \textit{Curation} is addressed as well, because data can be referenced with precision and maintenance of every data can be done in one place. 
+Since the use of RDF allows us to add an additional link between ID and value, we can separate the two and disclose all information about the data in very detail in machine-understandable form, and yet protect the values (\textit{Protection}).
+Furthermore, when every value can be referred to from anywhere, a simple technique can be applied to enable format-independent reconstruction of data files for which we introduced the concept of a \textit{data template}.
+Particularly the document-specific declaration of tokens enables a format-independent solution (\textit{Formats}) as this flexibility allows us to avoid syntactical clashes with the particular data format used.
+We have used SPARQL endpoints as a generic implementation of the \textit{API} which only served for the purpose of data retrieval.
+For actual application, a more elaborated API should be developed that could be based on Web Services.
+
+Eventually, we have to show that the restored files equal their originals on character level (\textit{Equality}).
 If the two expressions
@@
 The ideas presented in this work can be used to offer data and code behind research publications, which increases transparency, re-use, and reproducibility.
-
-We have formulated requirements
-
-We have demonstrated how fine grained data referencing in our data restore model can overcome the problem of data protection.
+We have formulated requirements on the basis of the recommendations formulated in \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337} and further followed the idea of grained data referencing in our data restore model.
 %Within the data template single data values can be addressed, even if the original data set is not publicly available.
-The introduced data templates provide precise pointers to the data used and can be disseminated freely as they do not disclose sensitive content.
+The introduced concept of data templates uses precise pointers to the data used and can be disseminated freely as they do not disclose sensitive content.
 Since data files can be restored in the format used by the researcher, the template technique closes the gaps between data citation and program code integration.
+The \textit{ddocks} prototype illustrates how the restore model can be applied in a way that enables a "one-click" solution for re-running analyses on other systems provided that users have permission to acquire the data and system dependencies are resolved.
+Especially If linked with research articles, the resources can be used in tutorials to teach computational research in practice, serve as re-usable research resources, or can be used within review-processes.
 With the precise identification of values, provenance of the data used can be clarified comprehensively whereas responsibilities for maintenance and documentation of a data set are left to one archive respectively, so that harvesting techniques can be used efficiently to carry forward documentation, notes and updates about the data.
-
-%Researchers can make use of these information resources directly from within their research environment and run on updated data....
-
-We have overcome some of the obstacles (data protection, distributed sources, tracability, source code integration vs.
-citation) and shown how a "one-click solution" to reproduce results can be implemented in which the gaps between data citation and code integration are closed.
 
@@
 While most statistical models today exist on paper only, the current work on reproducibility allows to rerun it again on the original or updated data, and the vision is to make them re-usable for other data as well.
-A research publication could remain more relevant for extended periods of time if a prognostic model on time series data for example can still be used on new records while the published article is no longer relevant to the community.
+A research publication could remain more relevant for extended periods of time if its prognostic model on time series data for example can still be used on new records while the published article may no longer be relevant to the community.
 Linked with thesauri, researchers can find these models by topic and re-use them in their own context while valuable user assistance can be given automatically if the semantic annotations specify clearly the kind of data it can process meaningfully.
 
-Moreover, these models could serve as input for the artificial intelligence community as they often describe abstract concepts like "living standards" or "justice" which are typically hard to model.
+Moreover, these models could serve as input for the artificial intelligence community as they often describe measures for abstract concepts like \textit{living standards} or \textit{justice} which are typically hard to model.
 Although these models are and should always be debatable, the semantic linkage to statistical indicators gives information on aspects that are relevant for understanding these concepts.
@@
 
-%
-% Vision
-%
-
-The material can be used in tutorials to teach computational research in practice, serve as re-usable research resources, or can be used within review-processes.
-
-%BZ: was genau meinst du mit the material? unseren code? restored data (references)?
-
-Considering a tight connection of data or data references with code and the eventually text of a research publication, the idea of a data and code ontology comes up.
-
-Such an ontology could be used to model the interconnection between data, code and research results in detail.
+Considering a tight connection of data or data references with code and the eventually text of a research publication, the idea of a data and code ontology comes up that models the interconnection between data, code and research results in detail.
 Moreover, if also containing the data values of our data template, it could also be used as an intermediate data set for reproduce data computations without having access to the original data.
-
-We will investigate the design and development of a data and code ontology in our future work.
-
-The ideas presented in this work could be used to offer data and code behind statistical research publications to students and researchers.
-We have demonstrated how detailed referencing can be used in a data restore model that 
- * overcome problem of data protection
-
-enables the
-
-The material could be used in tutorials to teach computational research in practice, serve as re-usable research resources, or could be used within review-processes.
-
-or as part of review-pr
-
-Tutorials could be 
-The material could help learn and understand research in practice
-
-Scalability needs to be tested. If the idea is rolled out to the archives, would the servers with billions of IDs scale?
-
-%
-% So eine Ontologie könnte es erlauben Daten und Code so zu verweben, dass man gut anbieten kann anstelle der geschützten Daten einen offenen Zwischendatensatz zu verwenden, mit dem weitergerechnet werden kann oder so....
-%
-
-%
-% Input für eine Data and Code Ontologie
-%
-
-%
-% Anforderungen aus Economics und Social Sciences zur Verbesserung der Reproduzierbarkeit
-%
-% Code ist immer erwähnt. und sie sagen, wie die readme datei aussehen muss usw.
-% Diese ganzen elemente können in einer ontologie organisiert werden 
-%
-
-there are other recommendations in \cite{kleiber2012reproducible}   % replication files, code, technical information
-that give requirements for an ontology that organizes the data and code resources.
-
+We want to explore this scenario on the basis of \cite{arai2011fragile} for which Sweave sources are available while the data set is access-restricted.
+Yet, scalability remains to be tested in regard to curation efforts and technical performance.
+We will investigate the design and development of a data and code ontology in our future work for which we will take into account recommendation \textit{(C)} in \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337} and other
 %
@@
 \cite{KoenkerZeileis2009} 
+% haben wir oben schon erwähnt
 % mccullough, got replicability
-\cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337}
-
+%\cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337}
+%
 % further guidelines
 \cite{baiocchi2007reproducible}
-
+%
 % Gentleman: "important, if not essential, to integrate the computations and code used in data analyses,...."
 \cite{gentleman2007statistical}
-
+%
 % Reporting guidelines for simulation-based research in social sciences
 \cite{rahmandad2012reporting}
+for the purpose of replicability in computational research.
+% Replication attempts and intermediate results should be incorporated in an ontology as well
+% these outcomes can be linked and support the linked science ontology etc.
+% VCR, point (J) in mccullough
 
@@
 
+%
+%
+% cleanups for a camera-ready version:
+%
+% * intro, erster absatz: discipline-specific solutions are needed
+%
+% * outlook: data and code ontology, wie sieht es aus mit den Verifiable Computational Research resourcen, replication attempt und linked science ontology?
+%
+%
+%
+
 \balancecolumns

### 2013-05-06 paper-data-restore-model.Rnw
submitted version

@@
 On this basis, we develop the requirements for a data restore model and elaborate a generic design in view of an overall data management infrastructure.
-We further present a proof of concept implementation on the basis of a real-world example from economics.
+We further present an experimental implementation on the basis of a real-world example from economics.
 Eventually we close with the vision of a data and code ontology that carries statistical models from paper to a re-usable semantic level.
@@
 % panton principles
-\cite{murray2010panton}
+%\cite{murray2010panton}
 
@@
 
-\subsection{Requirements}
+\subsection{Requirements}\label{sec:req}
 
@@
 The number of IDs needed is certainly high and questions of scalability might rise immediately.
-To give an estimation on the technical implications, the number of additional triples needed is twice the number of data values, or the same number only if linking with other semantic resources is not made use of.
+To give an estimation on the technical implications, the number of additional triples needed is twice the number of data values, or the same number only if not linked with other semantic resources.
 Maintenance could be supported with tools and there should not be any additional efforts, provided that the data is maintained in a structured representation already.
@@
 However, we would like to have a discussion on this naïve approach and leave the investigation of scalability for now.
+This could also help solve some of the problems in using linked data for scientific purposes.
+\cite{bechhofer2010linked}\footnote{The problem of precise data references has been recognized by the Semantic Statistics community as well \url{http://datalift.org/en/event/semstats2013}.}
 
@@
 
-In case the researcher used own data, the data would have to be submitted to the dedicated archive first where the respective data template and its IDs involved are generated first.
+In case the researcher used own data, the data would have to be submitted to an affiliated archive\footnote{Researchers have to archive their data, either locally or at another archive.
+By "affiliated archive" we denote this repository.} first where the respective data template and its IDs involved are generated first.
 The archive returns the template to the researcher who can then publish it in the same way as described above.
@@
 Furthermore, these tools should provide for further semantic annotation of the content (using the RDF Data Cube Vocabulary for example) so that documentation about the data is available as well.
+Figure \ref{fig:modelprocessing} outlines the overall process.
 
-Figure \ref{fig:modelprocessing} outlines the overall process and suggests roles and responsibilities for archives and digital libraries.
-The data itself would be maintained at the archives where Web Services provide access to up-to-date information and, if the policies allow, to its contents.
+%Figure \ref{fig:modelprocessing} outlines the overall process and suggests roles and responsibilities for archives and digital libraries.
+%The data itself would be maintained at the archives where Web Services provide access to up-to-date information and, if the policies allow, to its contents.
 %Data templates and source code resources behind a scientific article could be maintained at the library
@@
 %\section{An example from the economics}\label{sec:impl}
-\section{Proof of Concept}\label{sec:impl}
+\section{Experimental Implementation}\label{sec:impl}
 
@@
 
+We validate this experimental implementation with respect to the requirements developed in Section \ref{sec:req}.
 As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set could be referenced from any context (\textit{Tracability}).

### 2013-05-07 paper-data-restore-model.Rnw
cleanups

@@
 
-\newcommand{\conceptual}[1]{
-%\textcolor{RoyalBlue}{#1}
-\texttt{#1}
-}
-
-% TODO: wir sollten noch die Terminologie klï¿½ren. wie bezeichnen wir 
-%
-% * das Prinzip, Werte zu referenzieren:  Bisher nannte ich es Reference Model
-%
-% * so ein Daten-Template, das IDs anstelle von Werten enthï¿½lt: data reference template? ddocks template?
-%
-%
-% Solange das noch nicht so ganz gefunden ist, kï¿½nnen wir einfach diese Tokens verwenden:
-%
-\newcommand{\refmodel}{
-%data reconstruction model
-data restore model
-}
 \newcommand{\datatemplate}{
@@
 \title{
-%On the Reproducibility of Statistical Research
-%On the Reproducibility in Computational Statistics
 A Data Restore Model\\
@@
 One goal is to enable data publication in difficult settings.
-%BZ: "this research" koennte verwechselt werden mit der research, die reproduzierbar sein soll. wie waers mit "our goal is to enable ..."
 Moreover, we show how the gaps between data citation and code integration can be closed as to eliminate all manual efforts of arranging code and data for reproduction attempts. 
@@
 
+\newpage
+
 \section{Introduction}
@@
 Experiments must lead to the same findings when conducted by other persons or in other environments, under the premise of keeping relevant and considered parameters constant.
-%This is very often not the case in computational research 
 % sciencemag article on "Reproducible research in computational science"
@@
 \cite{peng2011reproducible}
-%One of them is the plain fact that the data on which the research is built often remains unavailable to others which gives rise for a more organized data management practice. \cite{Wood1298248}
 One of them is the plain fact that the supporting data often is unavailable to others which gives rise for a more organized and open data management practice. \cite{Wood1298248}
@@
 And yet, there is more to be considered.
-%A comprehensive review of the JMCB\footnote{Journal of Money, Credit, and Banking} data+code archive reveals that results cannot be reproduced even when these resources are given.  
 A comprehensive review of the \textit{JMCB\footnote{Journal of Money, Credit, and Banking} data+code archive} reveals that provenance information for the data used is often incomplete and that submitted code does not reproduce the results either.
@@
 However, empirical research is often based on sensitive data that cannot be shared easily.
-%Such data sets can be transformed into anonymized public-use files in general which needs verification in terms of cell sizes and de-anonymization and requires additional efforts, often on the researchers' side.
 A general solution to this problem is to transform such data sets into anonymized public-use files that can be distributed without infringements of data protection.
@@
 The idea is to equip researchers with means that make full transparency possible even in difficult settings of protected and distributed data sources, so that reconstruction remains a question of access permission only.
-%We aim to design this referencing technique in a way that allows for seamless integration with preferred research tools and source code, so that no further manual work is required for setting up data and code resources during replication attempts.
-%The restore model is designed to seamlessly integrate with preferred research tools and source code, so that no further manual work is required for setting up data and code resources during replication attempts.
 The restore model is designed to bridge the gaps between data citation and program code and to seamlessly integrate with preferred research tools, so that no further manual work is required in setting up data and code resources for replication attempts.
@@
 
-%In this paper we aim to overcome these challenges for the domain of statistical data and elaborate on a holistic solution for an infrastructure that enables reconstruction of individual research data sets.
-%This infrastructure is basically realized by an underlying data reconstruction model, which builds the bridge between the researchers' preferred software tool and the research data. 
-%The approach follows the idea of a data reconstruction model by which we develop the requirements it carries on an infrastructure.
-%We present an implementation of the model that tries to seamlessly integrate with any software tool preferred by a researcher.
-%By utilizing the model generic ids are assigned to all data values regardless of their original format and their physical location.
-%This allows for reconstructing and referencing the data set out of the researchers programming tool.
+The remainder of the paper is structured as follows. 
+In Section \ref{sec:relatedwork} we discuss related work relevant to our approach. 
+%In section 2 we present a case study with real world examples, where literate programming with research data is applied. 
+%Based on these examples we propose a generic data reconstruction model in section 3. 
+The requirements for a data restore model are elaborated in Section \ref{sec:datareferencing} by which we develop a generic solution that.
+An experimental implementation is presented in Section \ref{sec:impl} where we explore its application in context of a case study from economics.
+%Finally, we conclude and present future work in Section \ref{sec:outlook}.
+Finally, we conclude in Section \ref{sec:outlook} with summary and outlook on future work and present a vision that illustrates the potential of this approach .
 
-The remainder of the paper is structured as follows. 
-In section 2 we present a case study with real world examples, where literate programming with research data is applied. 
-Based on these examples we propose a generic data reconstruction model in section 3. 
-We implement and validate our model in section 4. 
-In section 5 we discuss related work relevant to our approach. 
-Finally, we conclude and present future work in section 6.
-%BZ: veraltet. muss man am schluss aktualisieren, wenn das paper steht
-
-\section{Related Work}
+\section{Related Work}\label{sec:relatedwork}
 
@@
 In this section, we first develop the requirements for a data restore model.
-%Subsequently, we elaborate a generic design and outline the big picture of the overall data mangement infrastructure and workflows.
 Subsequently, we elaborate a generic design and discuss its implication on an overall data mangement infrastructure and the workflows involved.
@@
 
+In the following, we want to develop the ideas of a referencing technique and take into account the recommendations \textit{(A-J)} fromulated by \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337} which we refer to using angle brackets.
 Whichever data management solution is applied eventually, it should fit to the researchers' workflows and practices in order to be effective and accepted hence.
@@
 On the side of reuse, there should not be any efforts needed in composing and organizing these resources either, meaning that no puzzling and manual tinkering in organizing data and code to make it run is required \textit{(F partly)}.
-In the following, we want to develop the ideas of a referencing technique and take into account the recommendations \textit{(A-J)} fromulated by \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337} which we refer to using angle brackets.
 Location and structure of data files on the local file system should be inherently specified, so that it integrates perfectly with the code.
@@
 
-% zu weich, was heißt das?  
 %  \item overall, it has to meet the constraints and requirements from all parties involved in the overall data management infrastructure, such as archives, researchers and libraries.
@@
 
-%As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set could be referenced from any context (\textit{Tracability}).
-%The requirement item \textit{Curation} is addressed as well, because data can be referenced with precision and maintenance of every data can done in one place. 
-%Since the use of RDF allows us to add an additional link between ID and value, we can separate the two and disclose all information about the data in very detail in machine-understandable form, and yet protect the values (\textit{Protection}).
-%Furthermore, when every value can be referred to from anywhere, a simple technique can be applied to enable format-independent reconstruction of data files.
-%We therefore introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values which solves the requirements of \textit{Equality}.
-%The data template contains a header section in which we can specify encoding and tokens that are used to identify IDs within the template.
-%It enables the format-independent solution (\textit{Formats}) as this flexibility allows us to avoid syntactical clashes with the particular data format used.
-%In addition, we also declare Web Service addresses for every namespace used (\textit{API}) under the pragmatic assumption that every archive provides one such Web Service for data enquiries only.
-%At this point, we do not want to elaborate on the Web Service interface, however, we have used SPARQL endpoints in the proof of concept implementation presented in Section \ref{sec:impl}.
-
 \subsection{Workflows}
@@
 \caption{
-Using the \refmodel
+Using the data restore model
 }
@@
 \caption{
-Excerpt from one of the data templates
+%Excerpt from one of the data templates
+%Data template excerpt from one of the data templates
+Data template for one of the two data sets (\textit{rk.raw})
 }
@@
 %
+% link to source code
 %
-%\begin{quote}
-
 \url{https://github.com/dbahls/ddocks-utils.git}
-%\end{quote}
-
-%For a more detailed view on the implementation, please take a look at the source code.\footnote{\url{https://github.com/dbahls/ddocks} \url{}}
 
@@
 Researchers could embed these calls within their source code to load the data by template at the beginning of their analyses without being bothered, and no further modifications are needed at the time of publishing.
-%This allows us to comfortably use the code within the source code of a research project from the begin
+
 %
+% link to source code
 %
-%\begin{quote}
-
 \url{https://github.com/dbahls/ddocks.git}
-%\end{quote}
 
@@
 We also inserted the two restore calls above in the source code of our example paper (\cite{koenkerzeileis09}) right before the load-data calls respectively, and running Sweave on it produced a report identical to the published article.\footnote{using R x64 2.15.3 on Windows}
+
 %
+% link to source code
 %
-%\begin{quote}
+\url{https://github.com/dbahls/paper-restore-model.git}
 
-\url{https://github.com/dbahls/paper-restore-model.git}
-%\end{quote}
-
-\section{Summary, Vision and Outlook}
+\section{Summary, Vision and Outlook}\label{sec:outlook}
 

### 2013-05-07 paper-data-restore-model.tex
cleanups

@@
 \caption{
-Excerpt from one of the data templates
+%Excerpt from one of the data templates
+%Data template excerpt from one of the data templates
+Data template for one of the two data sets \textit{rk.raw}
 }

### 2013-05-07 paper-data-restore-model.Rnw
language

@@
 Researchers are more and more requested to publish their scientific data sets for purposes of transparency, re-use, and reproducibility.
-Particularly in economics and the social sciences, researchers often use sensitive statistical data that underlies protection policies that inhibit distribution to third party archives.
-In addition, a substantial quantity of data sets is combined from one or more external providers, which easily complicates curation-related activities. %---when working with data copies
-These circumstances give us reason to pursue a data restore model based on fine-grained referencing that allows to trace data provenance to the original archive in charge of curation.
-One goal is to enable data publication in difficult settings.
-Moreover, we show how the gaps between data citation and code integration can be closed as to eliminate all manual efforts of arranging code and data for reproduction attempts. 
-On this basis, we develop the requirements for a data restore model and elaborate a generic design in view of an overall data management infrastructure.
-We further present an experimental implementation on the basis of a real-world example from economics.
+Particularly in economics and the social sciences, researchers often use sensitive statistical data that underlie protection policies which inhibit distribution to third party archives.
+In addition, a considerable quantity of data sets combines data from one or more external providers, which complicates the setting for curation-related activities. %---when working with data copies
+These circumstances give us reason to pursue a \textit{data restore model} on the basis of fine-grained referencing that allows to trace data provenance to the original archive in charge of curation.
+%One goal is to enable data publication in difficult cases.
+%Moreover, we show how the gaps between data citation and code integration can be closed as to eliminate all manual efforts of arranging code and data for reproduction attempts. 
+One goal is to enable data publication in difficult cases, and another one is to show how the gaps between data citation and code integration can be closed in order to eliminate all manual efforts of arranging code and data files for reproduction attempts.
+On this basis we develop the requirements for a data restore model and elaborate a generic design in view of an overall data management infrastructure.
+We further explore an experimental implementation which we validate by taking the example of a real-world publication in economics.
 Eventually we close with the vision of a data and code ontology that carries statistical models from paper to a re-usable semantic level.
@@
 % sciencemag article on "Reproducible research in computational science"
-In computational research, this remains unverified in many cases due to various reasons.
-\cite{peng2011reproducible}
-One of them is the plain fact that the supporting data often is unavailable to others which gives rise for a more organized and open data management practice. \cite{Wood1298248}
+In computational research, this remains unverified in many cases due to various reasons \cite{peng2011reproducible}.
+One of them is the plain fact that the supporting data is often unavailable to others which gives rise for a more organized and open data management practice \cite{Wood1298248}.
 While a scientific publication is commonly understood as synonymous with a peer-reviewed article alone, program code and data take the role of nice-to-have additional material only that rarely find their way to the public.
@@
 shows that inconsistencies occur in the social sciences as well.
-The data was available to the community \cite{dataset_FNSEM} which allowed other researchers to investigate and after all heavily criticize their publication as its results could not be reproduced.
-\cite{arai2011fragile}
+The data was available to the community \cite{dataset_FNSEM} which allowed other researchers to investigate and after all heavily criticize their publication as its results could not be reproduced \cite{arai2011fragile}.
 To underscore their critique, they published the source code of their own analyses online and in this way called for clarification. % and revealed their difficulties in replicating.
 The original authors responded with an errata in which they acknowledged the faults.
-At the same time, they presented a modified version of the statistical model they had applied and managed to produce the previous findings again while admitting that ``the results are now less clear-cut and [they lost] statistical significance''.
-\cite{bisin2011errata}
+At the same time, they presented a modified version of the statistical model they had applied and managed to produce the previous findings again while admitting that ``the results are now less clear-cut and [they lost] statistical significance'' \cite{bisin2011errata}.
 %
@@
 "An article about computational science in a scientific publication is not the scholarship itself, it is merely advertising of  the scholarship.
-The actual scholarship is the complete software development environment and the complete set of instructions which generated the figures."
-\cite{Buckheit1995b}
-\cite{de2001reproducible}
+The actual scholarship is the complete software development environment and the complete set of instructions which generated the figures" \cite{Buckheit1995b} \cite{de2001reproducible}.
 In support of this view \cite{KoenkerZeileis2009} investigated reproducibility in econometrics and pointed out how software tools can be used to organize data, code, and documentation in this domain.
@@
 With the increasing use of statistics software \cite{RePEc_hum_wpaper_sfb649dp2011_020}
-these problems emerge in the domain of empirical research in general where more and more statisticians promote the use of literate programming techniques  
+these problems emerge in empirical research in general where more and more statisticians promote the use of literate programming techniques  
 % Executable Papers for the R Community: The R2 Platform for Reproducible Research
@@
 \cite{rossini2003literate}
-as to combine report text and coding in single executable scripts so that basic documentation of the analyses is already given naturally.
+%as to combine report text and coding in single executable scripts so that a basic documentation of the analyses is already given naturally.
+which combine report text and coding in single executable scripts in a way that clarifies the link between research code, the figures produced, and the findings naturally.
 In both of the examples above, the authors made use of the literate programming tool Sweave \cite{lmucs_papers_Leisch_2002} 
-so that the report can be generated again from data processing to generating the report document, under the premise that the data set is given and the system environment allows for correct execution.
+so that their reports can be generated again, from data processing to generating the report document, under the premise that the data is given and the system environment allows for correct execution.
 
-However, empirical research is often based on sensitive data that cannot be shared easily.
-A general solution to this problem is to transform such data sets into anonymized public-use files that can be distributed without infringements of data protection.
+%However, 
+Empirical research is often based on sensitive data that cannot be shared easily.
+%A general solution to this problem is to transform such data sets into anonymized public-use files that can be distributed without infringements of data protection.
+One solution to this problem is to transform such data into anonymized public-use files that can be distributed without infringements of data protection.
 But doing so requires careful verification in terms of cell sizes and de-anonymization and requires additional efforts, often on the researchers' side.
 Moreover, researchers hesitate to publish their own data if incentives are not clear, especially when they had a lot of work producing it.
-In addition, data is often obtained from several sources and the mere citation of them does not give any details on the file formats used or way of composition as used for input for the computer program.
+In addition, data is often obtained from several sources and the mere citation of them does not give any details on the file formats used or way of composition as used as input for the computer program.
 %BZ: hier kommst du wieder zurueck zu den daten (wenn auch mit anderem schwerpunkt), nachdem das thema oben eigentlich abgehakt wirkte und du zum thema program code und danach zu provenance gegangen bist.
@@
 %These circumstances motivate our efforts to elaborate on a holistic solution for the statistical domain that enables the exact reconstruction of individual research data sets as were used by the researchers.
-In this paper we aim to overcome these challenges for the domain of statistical data and elaborate on a holistic solution that enables the exact reconstruction of individual research data sets as were used by the researchers.
+In this paper we aim to overcome these challenges for the domain of statistical data and elaborate on a holistic solution to enable the exact reconstruction of individual research data sets as used originally by a researcher.
 We pursue the idea of a \emph{data restore model} that is based on fine-grained data referencing as described in earlier works \cite{BahlsTochtermann_2012}
 where further background and motivation is explained more thoroughly.
-As this referencing technique can be viewed as means for precise citation, the linkup to provenance and further documentation on the particular data version is given which again contributes to clarity.
+Since this referencing technique can be viewed as means for precise citation, the linkup to provenance and further documentation on the particular data version is given which again contributes to clarity.
 The idea is to equip researchers with means that make full transparency possible even in difficult settings of protected and distributed data sources, so that reconstruction remains a question of access permission only.
@@
 % da muss man mal sehen, ob das wirklich alles rein muss
-The problem of reproducibility in computational research is addressed by many.
-If re-execution of experiments is the goal, it is necessary to capture information on the entire system environment used as to make sure respective processes can be performed under the same parameters.
+The problem of reproducibility in computational research has been addressed by many.
+If re-execution of experiments is the goal, it is necessary to capture information on the entire system environment used in order to make sure respective processes can be performed under the same parameters
 %
@@
 \cite{Rechert_2010_EBS_1816123_1816182}
+.
 %\cite{Guttenbrunner_2012_MFE_2180868_2180876}
 %This alone is already a very hard problem, because it must include hardware and software dependencies to very detail.
-This alone is already a very hard problem, because the chain of dependencies in terms of hardware and software has to be captured and restored to the very detail.
+This alone is a very hard problem already, because the chain of dependencies in terms of hardware and software has to be captured and restored to the very detail.
 In this regard, \cite{Schwab_2000_MSC_369545_369555} illustrates how the problem of software dependencies for computational research could be solved with the help of build scripts such as make files.
@@
 %Nevertheless, the achievements in this area and the opportunities for application such as \cite{Nowakowski2011608} seem promising.
-As the goal of our research is to find means for precise reconstruction of data sets and direct integration with program code, these approaches form a good complement. % todo: formulierung   
+As the goal of our research is to find means for precise reconstruction of data sets and direct integration with program code, these approaches make a good complement. % todo: formulierung   
 %BZ: As we aim to find means for precise reconstruction of data sets and direct integration with program code, the approaches complement each other.
 
-Particularly for the domain of empirical research, we found the literate programming approach very appealing, as it seems light-weight and suitable for less complex scenarios that can be handled with a single script.
+Particularly for the domain of empirical research, we found the literate programming approach very appealing as it seems light-weight and suitable for the many, less complex scenarios.
 It interweaves source code and text for the report, so that all processing steps are already quite well-documented. 
@@
 
-Source code for computational research does not need to be executable in order to make a valuable resource for transparency,
-as it is a detailed documentation of the processing already and a valuable resource as such, which is part of the argumentation in 
+Source code for computational research does not need to be executable in order to make a valuable contribution to transparency
+as it is a detailed documentation of the processing and a valuable resource as such, which is part of the argumentation in 
 % all these resources should have an identifier, as the title suggests
@@
 \cite{gavish2011universal}.
-The Verifiable Computational Research community aims to collect data, code and intermediate results of research processes as a trace to make the findings verifiable.
-Their automatic capturing process is integrated with the researchers' workflow, and therefore successful as well as unsuccessful attempts can be reviewed eventually. 
+The Verifiable Computational Research community aims to collect data, code and intermediate results of research processes as a trace to make findings verifiable.
+Their automatic capturing process is integrated with the researchers' workflow, and hence successful as well as unsuccessful attempts can be reviewed eventually. 
 %However, the VCR community argues that the source code does not have to be executable, source code as a detailed documentation of the processing suffices, and other intermediate results should be captured, so that the research is tracable and verifiable.
@@
 %BZ: Ich wurde den letzten teil ab "if it is" einfach streichen...
-Either way, source code should be picked up as part of the publication material whenever possible.
+Either way, source code should be picked up as part of the publication material whenever possible
 %
@@
 % science code manifesto
-\cite{barnes2012science}\footnote{\url{http://sciencecodemanifesto.org/}}
+\cite{barnes2012science}\footnote{\url{http://sciencecodemanifesto.org/}}.
 
@@
 
-In this section, we first develop the requirements for a data restore model.
+We begin this section by developing the requirements for a data restore model.
 Subsequently, we elaborate a generic design and discuss its implication on an overall data mangement infrastructure and the workflows involved.
@@
 
-In the following, we want to develop the ideas of a referencing technique and take into account the recommendations \textit{(A-J)} fromulated by \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337} which we refer to using angle brackets.
-Whichever data management solution is applied eventually, it should fit to the researchers' workflows and practices in order to be effective and accepted hence.
-\cite{Feijen_2011} 
+In the following, we want to develop the ideas of a referencing technique and take into account the recommendations fromulated by \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337} which we refer to using parentheses \textit{(A-J)}.
+Whichever data management solution is applied eventually should fit to the researchers' workflows and practices in order to be effective and thus accepted \cite{Feijen_2011}.
 In the ideal case, researchers can use their preferred tools with no further hassle preparing anything at the time of data submission.
@@
 %On the side of reuse, there should not be any unnecessary efforts in re-using these resources either, meaning that no puzzling and manual tinkering in organizing data and code to make it run.
-On the side of reuse, there should not be any efforts needed in composing and organizing these resources either, meaning that no puzzling and manual tinkering in organizing data and code to make it run is required \textit{(F partly)}.
+On the side of reuse, there should not be any efforts in composing and organizing these resources either, meaning that no puzzling and manual tinkering in organizing data and code is required to make it run \textit{(F partly)}.
 Location and structure of data files on the local file system should be inherently specified, so that it integrates perfectly with the code.
@@
 There are plenty of prevailing file formats, ranging from CSV, XML, Excel, entire RDBMS dumps to specific software-dependent and even unknown individual ``raw" types of formats.
-McCullough et al. recommend the use of ASCII encoding \textit{(D)} as to support re-usability, we therefore want to exclude binary formats from our discussion.
-As to be less restrictive and support more contemporary encodings such as UTF-8, we want to pursue the idea of using self-describing documents with information on its own encoding contained.
+McCullough et al. recommend the use of ASCII encoding \textit{(D)} for the benefit of re-usability, we therefore want to exclude binary formats from our discussion.
+As to be less restrictive and support more contemporary encodings such as UTF-8, we want to pursue the idea of self-describing documents with information on its own encoding contained.
 \footnote{Otherwise, data would need to be converted back and forth which causes trouble whenever a character has no representation in the ASCII standard}
-However, a format-independent solution should be found whereas most importantly, any restored data set must equal the original as to perfectly integrate with the code in the end \textit{(E)}. 
+A format-independent solution should be found whereas most importantly, any restored data set must equal the original as to perfectly integrate with the code in the end \textit{(E)}. 
 % which gives us a clear (requirement for successful) test case
-Thus, we note that data citation by persistent identifier (as is the current practice in data publishing) is not sufficient in this regard, since these identifiers point to data sets on a more abstract level only and do not distinguish or specify a particular format.
+Thus, we note that data citation by persistent identifier (as is the current practice in data publishing) is not sufficient for this purpose, since these identifiers point to data sets on a more abstract level only and do not distinguish or specify a particular format.
 \footnote{Most data agencies offer download in various formats while the content is filed under the same persistent identifier.}
@@
 
-Either way, pointers to the provenance of the data used should be available that give documentation and further background such as contact information to creators and curators for any version being used.
+Either way, pointers to the provenance of data should be available that give documentation and further background such as contact information for any version being used.
 While new persistent identifiers are commonly given for major data updates and releases only, those versions with minor bug-fixes cannot be referred to with precision so far.\footnote{Surely, an author can give a note on the time of data download or acquisition in the paper.
@@
 
-In contrast to the copy-based approach, the referencing technique allows data and documentation to be maintained at one responsible archive only, so that no duplicate efforts have to be made regarding data curation which is particularly relevant in cases where several external data sources were used.
-This improves quality of service as all enquiries about the data can be answered with the highest possible level of expertise available and gives a clear view on responsibilities in addition.
+In contrast to the copy-based approach, the referencing technique allows data and documentation to be maintained at one responsible archive only, with the result that no duplicate efforts have to be made in terms of data curation which is particularly relevant in cases where several external sources were used.
+This improves quality of service as all enquiries about data can be answered with the highest possible level of expertise available and gives a clear view on responsibilities in addition.
 
-As a result from this discussion we formulate the following requirements for the data restore model:
+As a result of this discussion we formulate the following requirements for the data restore model:
 
@@
 
-  \item \textbf{Curation:} The technique makes possible that data could be maintained and curated at those archives only that are responsible of maintenance with the required level of expertise.
+%  \item \textbf{Curation:} The technique makes possible that data could be maintained and curated at those archives only that are responsible of maintenance with the required level of expertise.
+%  \item \textbf{Curation:} The technique makes possible that data maintenance and curation needs to be done for the original copy only, right at the responsible archive, whereas harvesting techniques work efficiently
+  \item \textbf{Curation:} Using the model, data maintenance and curation needs to be done for the original copy only, right at the responsible archive, whereas documentation can be distributed efficiently on the basis of harvesting techniques.
 
@@
 %  \item Researchers are not disturbed by the introduction of technical hinderances such as new data formats or other obligations.
-  \item \textbf{Usability:} The method should integrate seamlessly with the researchers' workflows and suit to their preferred data formats. 
+%  \item \textbf{Usability:} The method should integrate seamlessly with the researchers' workflows and suit to their preferred data formats. 
+  \item \textbf{Usability:} The method should integrate seamlessly with the researchers' workflows and comply with their preferred data formats. 
 		% It should be independent of any particular data format. % kann man das so reinschreiben, klingt jedenfalls komisch
@@
 %  \item It should be simple and transparent, so that problems can be traced easily. % mit hilfe der direkten pointer ist alles gestochen scharf und einfach zu debuggen (?)
-  \item \textbf{Tracability:} The technique itself must be transparent, so that problems can be traced and questions can be answered efficiently.
+%  \item \textbf{Tracability:} The technique itself must be transparent, so that problems can be traced and questions can be answered efficiently.
+  \item \textbf{Traceability:} The technique itself must be traceable, so that references to data content can be followed efficiently, by humans and machines.
  %wenn du mit traceable meinst, dass zb berechnungen auf den daten nachvollziehbar sein sollen, dann wuerde ich hier auf jeden fall die ids nennen. transparent ist hier dann als begriff zu allgemein.
@@
 However, format-independent reconstruction of individual data files requires a mechanism that renders data content into any unknown or unspecified structure.
-The referencing technique implies that data slices of larger data sets can be identified and selected precisely, from global scope.
+The referencing technique implies that data slices of larger data sets can be identified and selected with precision, from global scope.
 Pointers to respective repositories in combination with SQL, SPARQL or other kinds of queries could be a solution.
-A challenge in this approach is that such queries produce result sets in the shape of tables which yet need to be transformed to this unspecified format.
+A challenge in this approach is that such queries produce result sets in the shape of tables which yet need to be transformed into this unspecified format.
 Whereas this idea is certainly worth exploring, we want to pursue another approach that fits better into the Semantic Web scheme and recent developments in the Linked Open Data movement.
@@
 %BZ: Hier suche ich noch ein, zwei Referenzen raus, die zusammen mit Halb_LDOW08 aufgezaehlt werden koennen.
-These technologies break down the separation between black box data bundles and attached metadata as it interweaves data with its documentation.
+These technologies break down the separation between black box data bundles and attached metadata as they interweave data with its documentation.
 %BZ: These technologies break not only down the separation between black box data bundles (the raw data) and attached metadata as it interweaves data with its documentation.
 %BZ: They also enable the publication, sharing and interlinking of data.
-The data content is more accessible and interpretable to machines which sets the stage for a variety of new applications, e.g. for data retrieval. \cite{BahlsTochtermann_2012}
+The data content is more accessible and interpretable to machines which sets the stage for a variety of new applications, e.g. for data retrieval \cite{BahlsTochtermann_2012}.
 %BZ: das wesentliche fuer unseren approach ist hier eigentlich gesagt, alles andere wuerde wohl eher in die related work passen, oder? wuerde diesen absatz mit dem folgenden zusammenfassen.
@@
 However, we would like to have a discussion on this naïve approach and leave the investigation of scalability for now.
-This could also help solve some of the problems in using linked data for scientific purposes.
-\cite{bechhofer2010linked}\footnote{The problem of precise data references has been recognized by the Semantic Statistics community as well \url{http://datalift.org/en/event/semstats2013}.}
+
+We introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values.
+This could also help solve some of the problems in using linked data for scientific purposes \cite{bechhofer2010linked}\footnote{The problem of precise data references has been recognized by the Semantic Statistics community as well \url{http://datalift.org/en/event/semstats2013}.}.
 
@@
 
-If implemented in big scale as a holistic information infrastructure for statistical research data management, the presented approach has further implications on workflows and gives ideas for a big picture.
+If implemented in big scale as a holistic information infrastructure for research data management in statistics, the presented approach has further implications on workflows and gives ideas for a big picture.
 Data archives could offer the download of data templates along with the actual data, so that researchers have the precise references right from the beginning of their research.
-Alternatively, researchers could pull the data on the basis of such data template from within their research tool, provided that a respective plugin is available.
-The advantage in the latter case is that this procedure could be embedded within their program code, so that no further modifications of data and code need to be done at the time of publishing.
-The researcher simply submits code and data templates used, and other researchers can rerun the same analysis without any manual arrangements or collection of resources, simply by executing the code.
+Alternatively, researchers could pull the data on the basis of these data templates alone from within their research tool, provided that a respective plugin is available.
+The advantage in the latter case is that this procedure could be embedded within their program code and no further modifications on data and code need to be done at the time of publishing.
+The researcher simply submits code and data templates, and other researchers can rerun the same analysis without any manual arrangements or collection of resources, simply by executing the code.
 
 In case the researcher used own data, the data would have to be submitted to an affiliated archive\footnote{Researchers have to archive their data, either locally or at another archive.
-By "affiliated archive" we denote this repository.} first where the respective data template and its IDs involved are generated first.
+By "affiliated archive" we denote this repository.} in beforehand where the respective data template and IDs involved are generated first.
 The archive returns the template to the researcher who can then publish it in the same way as described above.
-This procedure must be supported with tools for mapping and generation of IDs.
+This procedure must be supported with tools for mapping and ID generation.
 Furthermore, these tools should provide for further semantic annotation of the content (using the RDF Data Cube Vocabulary for example) so that documentation about the data is available as well.
@@
 
-We introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values which solves the requirements of \textit{Equality}.
+%We introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values which solves the requirements of \textit{Equality}.
 The data template contains a header section in which we can specify encoding and tokens that are used to identify IDs within the template.
@@
 We validate this experimental implementation with respect to the requirements developed in Section \ref{sec:req}.
-As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set could be referenced from any context (\textit{Tracability}).
+As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set could be referenced from any context (\textit{Traceability}).
 The requirement item \textit{Curation} is addressed as well, because data can be referenced with precision and maintenance of every data can be done in one place. 

### 2013-05-07 paper-data-restore-model.tex
language

@@
 %Data template excerpt from one of the data templates
-Data template for one of the two data sets \textit{rk.raw}
+Data template for one of the two data sets (\textit{rk.raw})
 }

### 2013-05-08 paper-data-restore-model.Rnw
language

@@
 
-For further details, the reader is invited to take a look at the source code we published on GitHub.
-Respective links are given in the subsections
-where we introduce the concept of a data template using two data sets from the sources of our example paper \cite{KoenkerZeileis2009}. 
-The \textit{ddocks} plugin and a basic validation which are part of the Sweave sources for this article.
+In the following, we explain the concept of data templates by taking the example of two data sets from the sources of our example paper \cite{KoenkerZeileis2009}. 
+%The \textit{ddocks} plugin and a basic validation which are part of the Sweave sources for this article.
+A description of the \textit{ddocks} plugin implementation follows thereafter,
+and a basic validation is given at the end of this section which we have incorporated as part of the Sweave sources for the article the reader is currently looking at.
+The source code are available on GitHub, and respective links are given at the bottom of each subsection.
 
@@
 
-We also inserted the two restore calls above in the source code of our example paper (\cite{koenkerzeileis09}) right before the load-data calls respectively, and running Sweave on it produced a report identical to the published article.\footnote{using R x64 2.15.3 on Windows}
+We also inserted the two restore calls above in the source code of our example paper (\cite{KoenkerZeileis2009}) right before the load-data calls respectively, and running Sweave on it produced a report identical to the published article.\footnote{using R x64 2.15.3 on Windows}
 
@@
 The ideas presented in this work can be used to offer data and code behind research publications, which increases transparency, re-use, and reproducibility.
-We have formulated requirements on the basis of the recommendations formulated in \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337} and further followed the idea of grained data referencing in our data restore model.
+We have formulated requirements on the basis of the recommendations in \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337} and further followed the idea of fine-grained data referencing in our data restore model.
 %Within the data template single data values can be addressed, even if the original data set is not publicly available.
-The introduced concept of data templates uses precise pointers to the data used and can be disseminated freely as they do not disclose sensitive content.
+The introduced concept of data templates uses precise pointers to the original data sources and can be disseminated freely as they do not disclose sensitive content.
 Since data files can be restored in the format used by the researcher, the template technique closes the gaps between data citation and program code integration.
 The \textit{ddocks} prototype illustrates how the restore model can be applied in a way that enables a "one-click" solution for re-running analyses on other systems provided that users have permission to acquire the data and system dependencies are resolved.
-Especially If linked with research articles, the resources can be used in tutorials to teach computational research in practice, serve as re-usable research resources, or can be used within review-processes.
-With the precise identification of values, provenance of the data used can be clarified comprehensively whereas responsibilities for maintenance and documentation of a data set are left to one archive respectively, so that harvesting techniques can be used efficiently to carry forward documentation, notes and updates about the data.
+Especially if linked with research articles, the resources can be used in tutorials to teach computational research in practice, serve as re-usable research resources, or can be used within review-processes.
+With the precise identification of values, data provenance can be clarified comprehensively whereas responsibilities for maintenance and documentation of a data set are left to one archive respectively, so that harvesting techniques can be used efficiently to carry forward documentation, notes and updates about the data.
 
@@
 A research publication could remain more relevant for extended periods of time if its prognostic model on time series data for example can still be used on new records while the published article may no longer be relevant to the community.
-Linked with thesauri, researchers can find these models by topic and re-use them in their own context while valuable user assistance can be given automatically if the semantic annotations specify clearly the kind of data it can process meaningfully.
+Linked with thesauri, researchers could find these models by topic and re-use them in their own context while valuable user assistance can be given automatically if the semantic annotations specify clearly the kind of data it can process meaningfully.
 
@@
 %
-%
+% * Bibliography: unvollständige Angaben: De Leeuw, Barnes
 %

### 2013-05-08 paper-data-restore-model.Rnw
language

@@
 We further explore an experimental implementation which we validate by taking the example of a real-world publication in economics.
-Eventually we close with the vision of a data and code ontology that carries statistical models from paper to a re-usable semantic level.
+Eventually we close with the vision of a \textit{data and code ontology} that carries statistical models from paper to a re-usable semantic level.
 
@@
 And yet, there is more to be considered.
-A comprehensive review of the \textit{JMCB\footnote{Journal of Money, Credit, and Banking} data+code archive} reveals that provenance information for the data used is often incomplete and that submitted code does not reproduce the results either.
+A comprehensive review of the \textit{JMCB\footnote{Journal of Money, Credit, and Banking} data+code archive} revealed that provenance information for the data used is often incomplete and that submitted code does not reproduce the results either.
 It is not uncommon that even the original authors have difficulties understanding the details of their own work only few months after publishing already.
@@
 %Based on these examples we propose a generic data reconstruction model in section 3. 
-The requirements for a data restore model are elaborated in Section \ref{sec:datareferencing} by which we develop a generic solution that.
+The requirements for a data restore model are elaborated in Section \ref{sec:datareferencing} by which we develop a generic solution.
 An experimental implementation is presented in Section \ref{sec:impl} where we explore its application in context of a case study from economics.
@@
 McCullough et al. recommend the use of ASCII encoding \textit{(D)} for the benefit of re-usability, we therefore want to exclude binary formats from our discussion.
-As to be less restrictive and support more contemporary encodings such as UTF-8, we want to pursue the idea of self-describing documents with information on its own encoding contained.
-\footnote{Otherwise, data would need to be converted back and forth which causes trouble whenever a character has no representation in the ASCII standard}
+As to be less restrictive and support more contemporary encodings such as UTF-8, we want to pursue the idea of self-describing documents with information on its own encoding contained.\footnote{Otherwise, data would need to be converted back and forth which causes trouble whenever a character has no representation in the ASCII standard}
 A format-independent solution should be found whereas most importantly, any restored data set must equal the original as to perfectly integrate with the code in the end \textit{(E)}. 
@@
 This might sound surprising to the reader as sensitive data cannot be shared per se and a restore concept appears redundant.
-Nevertheless, we want to make the point that the spectrum from open to protected data is gradual, since some data is access-restricted only and can be obtained under certain policies and regulations.
+Nevertheless, we want to make the point that the spectrum from open to closed data is gradual, since some data is access-restricted only and can be obtained under certain policies and regulations.
 Whether or not a data set can be handed out should be a question on legal level only, the restore technique itself, however, should cater for these cases as to give means for repeatability in every scenario.
@@
 
-In contrast to the copy-based approach, the referencing technique allows data and documentation to be maintained at one responsible archive only, with the result that no duplicate efforts have to be made in terms of data curation which is particularly relevant in cases where several external sources were used.
+In contrast to the copy-based approach, the referencing technique allows data and documentation to be maintained at one responsible archive only, with the result that no duplicate efforts have to be made in terms of data curation which is particularly relevant in cases where (several) external sources were used.
 This improves quality of service as all enquiries about data can be answered with the highest possible level of expertise available and gives a clear view on responsibilities in addition.
@@
 %  \item \textbf{Usability:} The method should integrate seamlessly with the researchers' workflows and suit to their preferred data formats. 
-  \item \textbf{Usability:} The method should integrate seamlessly with the researchers' workflows and comply with their preferred data formats. 
+%  \item \textbf{Usability:} The method should integrate seamlessly with the researchers' workflows and comply with their preferred data formats. 
+  \item \textbf{Usability:} The method should integrate seamlessly with the researchers' workflows and comply with their preferred tools. 
 		% It should be independent of any particular data format. % kann man das so reinschreiben, klingt jedenfalls komisch
@@
   % auch soll es möglichst einfach auf neue unbekannte Datenformate anwendbar sein, mitunter um zukunftsfähig zu sein
-  \item \textbf{Formats:} The method should be sustainable and applicable to alternative, unknown or future data formats.
+  \item \textbf{Formats:} The method should be sustainable and applicable to common, alternative, unknown and future data formats.
 
@@
 A challenge in this approach is that such queries produce result sets in the shape of tables which yet need to be transformed into this unspecified format.
-Whereas this idea is certainly worth exploring, we want to pursue another approach that fits better into the Semantic Web scheme and recent developments in the Linked Open Data movement.
+In this work, we want to pursue another approach that fits better into the Semantic Web scheme and recent developments in the Linked Open Data movement.
 
@@
 
+\newpage
+
 %\section{A case study - Two real world examples}
@@
 %We have implemented
-In this section, we present a prototype implementation of the data restore model and a plugin called \textit{ddocks}\footnote{\textit{ddocks} is short for data docks which is meant to reflect the ``transshipment" character of the interface between data retrieval and processing.}
+In this section, we present a prototype implementation of the data restore model and a plugin called \textit{ddocks}\footnote{\textit{ddocks} is short for \textit{data docks} which is meant to reflect the ``transshipment" character of the interface between data retrieval and processing.}
 for the computing environment R.
@@
 
-In the following, we explain the concept of data templates by taking the example of two data sets from the sources of our example paper \cite{KoenkerZeileis2009}. 
+In the following, we explain the concept of data templates by taking the example of two data sets from the sources of \cite{KoenkerZeileis2009}. 
 %The \textit{ddocks} plugin and a basic validation which are part of the Sweave sources for this article.
@@
 and a basic validation is given at the end of this section which we have incorporated as part of the Sweave sources for the article the reader is currently looking at.
-The source code are available on GitHub, and respective links are given at the bottom of each subsection.
+The source code is available on GitHub, and respective links are given at the bottom of each subsection.
 
@@
 %Researchers could embed these calls within their source code to load the data by template at the beginning of their analyses and leave it there without being bothered.
-Researchers could embed these calls within their source code to load the data by template at the beginning of their analyses without being bothered, and no further modifications are needed at the time of publishing.
+Researchers could embed these calls within their source code to load the data by template at the beginning of their analyses without being bothered again, and no further modifications are needed at the time of publishing.
 
@@
 We validate this experimental implementation with respect to the requirements developed in Section \ref{sec:req}.
-As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set could be referenced from any context (\textit{Traceability}).
+As this approach is based on URIs, namespaces can be used to identify archives, and in combination with locally organized IDs every single value of any data set can be referenced from any context (\textit{Traceability}).
 The requirement item \textit{Curation} is addressed as well, because data can be referenced with precision and maintenance of every data can be done in one place. 
@@
 While most statistical models today exist on paper only, the current work on reproducibility allows to rerun it again on the original or updated data, and the vision is to make them re-usable for other data as well.
-A research publication could remain more relevant for extended periods of time if its prognostic model on time series data for example can still be used on new records while the published article may no longer be relevant to the community.
+A research publication could remain more relevant for extended periods of time if its prognostic model on time series data for example can still be used on new records while the specific results of a published article may no longer be relevant to the community.
 Linked with thesauri, researchers could find these models by topic and re-use them in their own context while valuable user assistance can be given automatically if the semantic annotations specify clearly the kind of data it can process meaningfully.
 
-Moreover, these models could serve as input for the artificial intelligence community as they often describe measures for abstract concepts like \textit{living standards} or \textit{justice} which are typically hard to model.
+Moreover, these models could serve as an input for the artificial intelligence community as they often describe measures for abstract concepts like \textit{living standards} or \textit{justice} which are typically hard to model.
 Although these models are and should always be debatable, the semantic linkage to statistical indicators gives information on aspects that are relevant for understanding these concepts.
-Further, prognostic models could be combined in ensembles and trained with machine learning techniques, which might again serve the computational statistics community in their research.
+Further, prognostic models could be combined in ensembles trained with machine learning techniques to further increase accuracy, which might again serve the computational statistics community in their research.
 
-Considering a tight connection of data or data references with code and the eventually text of a research publication, the idea of a data and code ontology comes up that models the interconnection between data, code and research results in detail.
+Considering a tight connection of data or data references with code and the eventually text of a research publication, the idea of a \textit{data and code ontology} comes up that models the interconnection between data, code and research results in detail.
 Moreover, if also containing the data values of our data template, it could also be used as an intermediate data set for reproduce data computations without having access to the original data.

### 2013-06-24 paper-data-restore-model.Rnw
final version

@@
 % sciencemag article on "Reproducible research in computational science"
-In computational research, this remains unverified in many cases due to various reasons \cite{peng2011reproducible}.
+In computational research, this remains unverified in many cases due to various reasons \cite{Peng02122011}.
 One of them is the plain fact that the supporting data is often unavailable to others which gives rise for a more organized and open data management practice \cite{Wood1298248}.
@@
 And yet, there is more to be considered.
-A comprehensive review of the \textit{JMCB\footnote{Journal of Money, Credit, and Banking} data+code archive} revealed that provenance information for the data used is often incomplete and that submitted code does not reproduce the results either.
+A comprehensive review of the \textit{JMCB\footnote{Journal of Money, Credit, and Banking} data+code archive} revealed that provenance information for the data used is often incomplete and that submitted code does not reproduce the results either \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337}.
 It is not uncommon that even the original authors have difficulties understanding the details of their own work only few months after publishing already.
@@
 The requirements can be realized in several ways.
-However, format-independent reconstruction of individual data files requires a mechanism that renders data content into any unknown or unspecified structure.
-The referencing technique implies that data slices of larger data sets can be identified and selected with precision, from global scope.
-Pointers to respective repositories in combination with SQL, SPARQL or other kinds of queries could be a solution.
-A challenge in this approach is that such queries produce result sets in the shape of tables which yet need to be transformed into this unspecified format.
-In this work, we want to pursue another approach that fits better into the Semantic Web scheme and recent developments in the Linked Open Data movement.
+However, giving freedom to researchers to use or reuse data in their preferred environments and associated formats needs careful consideration.
+Data could be offered in a variety of formats, each version provided with means for persistent identification, so that data citation as well as seamless code integration is ensured.
+But this approach requires program code to function on complete data sets, which impairs usability considering that researchers typically select slices of indicators and observations\footnote{e.g.
+selection of certain indicators and reference periods from large time series data} for input to their programs.
+Further, the approach would lead to multiple persistent identifiers for content-wise identical data and complicate bibliometric measures.
+These conditions suggest the use of a primary data representation from which all other formats can be derived whereas one referencing method only is needed to serve all scenarios.
+As a consequence, such model must give means for identification of any possible data slice from global scope.
+Pointers to respective repositories in combination with structured queries (such as SQL or SPARQL) could be a solution.
+A challenge in this approach, however, is that such queries produce result sets in the shape of tables which yet need to be transformed into this unspecified format as used by a researcher.
+In the following, we elaborate a model that aims to overcome these challenges.
 
-Currently, more and more data sets are released as Open Government data and statistical offices have already explored the use of semantic technologies as well \cite{Halb_LDOW08}.
+The model has to make sure that individual data slices and values can be identified without ambiguity in the long term.
+As we could not find an established data model for statistical data that is not subject to change and incorporates versioning at the same time, we discarded model-based queries and decided to use individual identifiers for every single value of a data set.
+The number of IDs needed is certainly high and questions of scalability raise immediately.
+On the other side, the IDs we are going to introduce can as well be interpreted as single-result queries or as the variables of a regular query.
+%To give an estimation on the technical implications, the number of additional triples needed is twice the number of data values, or the same number only if not linked with other semantic resources.
+%Maintenance could be supported with tools and there should not be any additional efforts, provided that the data is maintained in a structured representation already.
+%BZ: what tools?
+However, we would like to have a discussion on this na\"{\i}ve approach and leave the investigation of scalability for now.
+We will introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values.
+This could also help solve some of the problems in using linked data for scientific purposes \cite{bechhofer2010linked}\footnote{The problem of precise data references has been recognized by the Semantic Statistics community as well \url{http://datalift.org/en/event/semstats2013}.}, which will be the basis of the model.
+
+With the ongoing linked open data movement, awareness and knowledge about semantic technologies and the RDF data model has found its way to many application domains, among which the domain of statistical data\cite{Halb_LDOW08}.
 %BZ: Hier suche ich noch ein, zwei Referenzen raus, die zusammen mit Halb_LDOW08 aufgezaehlt werden koennen.
@@
 
-Our approach is based on the concept of assigning individual IDs to every single value of a data set.
-%BZ: Das ist notwendig, da man nicht das dataset als ganzen restoren moechte, sondern ja gerade die einzelnen data values benoetigt, um ggf.
-berechnungen nachvollziehen zu koennen.
-Bekomm das jetzt grad nicht griffig formuliert...
-The number of IDs needed is certainly high and questions of scalability might rise immediately.
-To give an estimation on the technical implications, the number of additional triples needed is twice the number of data values, or the same number only if not linked with other semantic resources.
-Maintenance could be supported with tools and there should not be any additional efforts, provided that the data is maintained in a structured representation already.
-%BZ: what tools?
-However, we would like to have a discussion on this naïve approach and leave the investigation of scalability for now.
+\subsection{Workflows}
 
-We introduce the concept of a \textit{data template} which is based on the idea to replace every value within a data file with its associated ID, so that the original file can be restored again simply by replacing them back with its associated values.
-This could also help solve some of the problems in using linked data for scientific purposes \cite{bechhofer2010linked}\footnote{The problem of precise data references has been recognized by the Semantic Statistics community as well \url{http://datalift.org/en/event/semstats2013}.}.
-
-\subsection{Workflows}
+% edawax survey sven vlaeminck
+%
+%\cite{LQURN:NBN:NL:UI:10-1-114595}
+%
 
@@
 
-\newpage
-
 %\section{A case study - Two real world examples}
@@
 The data template contains a header section in which we can specify encoding and tokens that are used to identify IDs within the template.
-In addition, we also declare Web Service addresses for every namespace used (which is only one for our example) under the pragmatic assumption that every archive provides one such Web Service for data enquiries only.
+In addition, we also declare Web Service addresses for every namespace used (which is only one for our example) under the pragmatic assumption that every namespace can be associated with one archive providing one such Web Service for data enquiries.
 %At this point, we do not want to elaborate on the Web Service interface, however, we have used SPARQL endpoints in the proof of concept implementation presented in Section \ref{sec:impl}.
@@
 While most statistical models today exist on paper only, the current work on reproducibility allows to rerun it again on the original or updated data, and the vision is to make them re-usable for other data as well.
-A research publication could remain more relevant for extended periods of time if its prognostic model on time series data for example can still be used on new records while the specific results of a published article may no longer be relevant to the community.
+A research publication could remain more relevant for extended periods of time if its predictive model on time series data for example can still be used on new records while the specific results of a published article may no longer be relevant to the community.
 Linked with thesauri, researchers could find these models by topic and re-use them in their own context while valuable user assistance can be given automatically if the semantic annotations specify clearly the kind of data it can process meaningfully.
@@
 Although these models are and should always be debatable, the semantic linkage to statistical indicators gives information on aspects that are relevant for understanding these concepts.
-Further, prognostic models could be combined in ensembles trained with machine learning techniques to further increase accuracy, which might again serve the computational statistics community in their research.
+Further, predictive models could be combined in ensembles trained with machine learning techniques to further increase accuracy, which might again serve the computational statistics community in their research.
 
-Considering a tight connection of data or data references with code and the eventually text of a research publication, the idea of a \textit{data and code ontology} comes up that models the interconnection between data, code and research results in detail.
-Moreover, if also containing the data values of our data template, it could also be used as an intermediate data set for reproduce data computations without having access to the original data.
+Considering the coupling of data or data references with code and the report text of a research publication eventually, the idea of a \textit{data and code ontology} comes up that models the interconnection between data, code and research results in detail.
+Moreover, it could also be used to model algorithmic transitions from original to intermediate and result data sets, which would help clarify data provenance and allow for reproduction from open intermediate data sets while original data might be closed an unaccessible.
 We want to explore this scenario on the basis of \cite{arai2011fragile} for which Sweave sources are available while the data set is access-restricted.
 Yet, scalability remains to be tested in regard to curation efforts and technical performance.
-We will investigate the design and development of a data and code ontology in our future work for which we will take into account recommendation \textit{(C)} in \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337} and other
+We will investigate the design and development of a data and code ontology in future work for which we take into account recommendation \textit{(C)} in \cite{RePEc_ejw_journl_v_4_y_2007_i_3_p_326_337} and other
 %
@@
 %
+% * id-approach: we should also mention: path-based approach: id-per observation + predicate-path to value, lower number of ids required, but maybe not precise enough?
+other difficulties then, but needs to be explored too (outlook/design)
+%
+%
 

### 2013-07-03 paper-data-restore-model.Rnw
camera-ready-version, i-know 2013

@@
 
-\documentclass{acm_proc_article-sp}
+%\documentclass{acm_proc_article-sp}
+\documentclass{sig-alternate}
 
@@
 
+\conferenceinfo{i-Know}{'13, September 04 - 06 2013, Graz, Austria}
+
+\CopyrightYear{2013}
+\crdata{978-1-4503-2300-0/13/09}
+
 \title{
@@
 }
+
 %
@@
 % A category with the (minimum) three required fields
+\category{H.3.5}{Online Information Services}{Data sharing}
 \category{H.3.7}{Digital Libraries}{Miscellaneous}
@@
 
-\terms{Design, Algorithms, Economics}
+%\category{Digital libraries and archives}
+
+\terms{Design, Economics, Algorithms, Verification}
+
+\newpage
 
@@
 
-\newpage
+%\newpage
 
@@
 The model has to make sure that individual data slices and values can be identified without ambiguity in the long term.
-As we could not find an established data model for statistical data that is not subject to change and incorporates versioning at the same time, we discarded model-based queries and decided to use individual identifiers for every single value of a data set.
+As we could not find an established data model for statistical data that is not subject to change and incorporates versioning at the same time, we discarded model-based queries and decided to use identifiers for every single value of a data set.
 The number of IDs needed is certainly high and questions of scalability raise immediately.
@@
 By "affiliated archive" we denote this repository.} in beforehand where the respective data template and IDs involved are generated first.
+At this point, metadata on provenance and other documentation should be collected which can be supported through scientific workflow systems\cite{cau21112}.
 The archive returns the template to the researcher who can then publish it in the same way as described above.
@@
 
+%\newpage
+
 \section{Summary, Vision and Outlook}\label{sec:outlook}
@@
 
-\balancecolumns
+%\balancecolumns
 
